---
title: "Understanding the variability in macromolecular composition of soil microbial biomass"
subtitle: "Supplementary Information"
author: "Petr Čapek, Michal Choma, Eva Kaštovská, Karolina Tahovská, Helen Glanville, Hana Šantrůčková"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and ggplot theme, echo=FALSE, warning=FALSE, message=FALSE}
#Libraries
library(ggplot2)
library(dplyr)
library(reshape)
library(minpack.lm)
library(bbmle)
library(deSolve)
library(FME)
library(DEoptim)
library(ABCoptim)
library(rcompanion)
library(optimx)
library(reticulate)
library(gridExtra)
#ggplot theme
theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
                 axis.title=element_text(size=18, colour="black"),
                 axis.line=element_line(size=0.5, colour="black"),
                 strip.text=element_text(size=18, face="bold"),
                 axis.ticks=element_line(size=1, colour="black"),
                 #axis.ticks.length=unit(-0.05, "cm"),
                 panel.background=element_rect(colour="black", fill="white"),
                 panel.grid=element_line(linetype=0),
                 legend.text=element_text(size=14, colour="black"),
                 legend.title=element_text(size=14, colour="black"),
                 legend.position=c("right"),
                 legend.key.size=unit(1, "cm"),
                 strip.background=element_rect(fill="grey98", colour="black"),
                 legend.key=element_rect(fill="white", size=1.2),
                 legend.spacing=unit(0.5, "cm"),
                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
```

## 1. Microbially-explicit models

### 1.1. Natural abundance of carbon isotopes

#### 1.1.1. Monod model

Monod model defines three pools of organic carbon (C) - organic substrate (S), microbial biomass (MB) and carbon dioxide ($CO_2$). The changes of these pools over time are defined by following equations:

$\frac{dS}{dt}=-uptake+death,$ (S1)

$\frac{dMB}{dt}=growth-death,$ (S2)

$\frac{dCO_2}{dt}=R_{growth}.$ (S3)

Microbial biomass consumes organic substrate ($uptake$) by a rate proportional to $MB$ and follows the hyperbolic scaling with $S$:

$uptake = I_m \times MB \times f;~~~f = \frac{S}{K_m + S}.$ (S4)

In eq. S4, $I_m$ is maximum uptake rate constant and $K_m$ is half-saturation constant. Organic substrate taken up is directly converted to $MB$ that has invariable chemical composition. The conversion of $S$ to $MB$ is denoted as $growth$. Growth requires energy. The energy requirements are covered directly from substrate and cause growth associated $CO_2$ production (i.e. $R_{growth}$). Growth and $R_{growth}$ are directly linked to $uptake$ via the parameter $y_A$, which defines the efficiency of organic carbon conversion to $MB$ (i.e. commonly denoted as carbon use efficiency):

$growth = uptake \times y_A,$ (S5)

$R_{growth} = uptake \times (1-y_A).$ (S6)

Along with the growth, $MB$ is constantly dying. The dying part of microbial community is consumed by the growing part of microbial community. The death rate follows the first order decay kinetic:

$death = MB \times d.$ (S7)

In eq. S7, $d$ is the death rate constant.

Since the $MB$ is not directly measured, invariable conversion factors between $MB$ and different proxy-parameters are defined as follows:

$CLC = MB \times k_{ec},$ (S8)

$DNA = MB \times k_{DNA},$ (S9)

$ATP = MB \times k_{ATP},$ (S10)

$PLFA = MB \times k_{PLFA},$ (S11)

$W = MB \times W_{MB}.$ (S12)

In eqs. S8 - S12, symbols $CLC$, $k_{ec}$, $k_{DNA}$, $k_{ATP}$, and $k_{PLFA}$ denote chloroform-labile organic carbon, and conversion factors between $CLC$, DNA, ATP and PLFA, respectively. Parameter $W_{MB}$ further denotes the C-1 mole to mass of total microbial community conversion factor.

#### 1.1.2. Pirt model

Pirt model is almost identical to Monod model. The only difference is the presence of maintenance. $MB$ has to maintain its integrity. Integrity maintenance requires energy, which is covered from $MB$ and results in $CO_2$ production ($R_{maintenance}$). $R_{maintenance}$ is proportional to $MB$ via the specific maintenance rate constant ($m$):

$R_{maintenance} = MB \times m.$ (S13)

The Pirt model is thus defined by following differential equations:

$\frac{dS}{dt}=-uptake+death,$ (S14)

$\frac{dMB}{dt}=growth-death-R_{maintenance},$ (S15)

$\frac{dCO_2}{dt}=R_{growth}+R_{maintenance}.$ (S16)

Conversion factors between different proxy-parameters and $MB$ is defined the same way as in Monod model (eqs. S8 - S12).

The Monod, Pirt and two-pool biomass model defined in main text represent gradient of increasing complexity and number of model parameters.

### 1.2. Isotope-specific models

#### 1.2.1. Sub-microbial model

The isotope-specific sub-microbial model have eight pools represented by symbols ${}^{12}S$, ${}^{12}B$, ${}^{12}G$, ${}^{12}CO_{2}$, ${}^{14}S$, ${}^{14}B$, ${}^{14}G$ and ${}^{14}CO_{2}$. The left superscripts of the symbols denote either natural carbon isotope 12 or radioactive isotope 14 (${}^{13}C$ labelled substrate is used instead of radioactive ${}^{14}C$ in study of Ziegler et al. (2005)). Sum of two isotopes in respective pool defines the total pool:

$S = {}^{12}S + {}^{14}S,$ (S17)

$B = {}^{12}B + {}^{14}B,$ (S18)

$G = {}^{12}G + {}^{14}G,$ (S19)

$CO_2 = {}^{12}CO_2 + {}^{14}CO_2.$ (S20)

The fluxes between the total pools are defined the same way as described in main text (section 2.1. Mathematical model). Isotope-specific fluxes are proportional to total fluxes by the relative abundance of the source pools. The relative abundance is defined in respect to radioactive carbon isotope as:

$S_{at} = \frac{{}^{14}S}{{}^{14}S + {}^{12}S},$ (S21)

$B_{at} = \frac{{}^{14}B}{{}^{14}B + {}^{12}B},$ (S22)

$G_{at} = \frac{{}^{14}G}{{}^{14}G + {}^{12}G},$ (S23)

$CO_{2at} = \frac{{}^{14}CO_2}{{}^{14}CO_2 + {}^{12}CO_2}.$ (S24)

Implementing eqs. S21 - S24 into eqs. 10 - 13 reported in main text, isotope-specific sub-microbial model is defined as:

$\frac{d^{12}S}{dt} = -U \times (1-S_{at}),$ (S25)

$\frac{d^{12}G}{dt} = A_m \times [f \times (1-S_{at})-\frac{G}{G_m}\times (1-G_{at})],$ (S26)

$\frac{d^{12}B}{dt} = \left\{ \begin{array}{rcl}\mu \times B \times (1-G_{at}) & \mbox{for}& \mu \geq 0 \\\ \mu \times B \times (1+g) \times (1-B_{at}) & \mbox{for} & \mu < 0\end{array}\right.,$ (S27)

$\frac{d^{12}CO_2}{dt} = \left\{ \begin{array}{rcl} R_{assimilatory} \times (1-S_{at}) + (R_{maintenance}+R_{growth}) \times (1-G_{at}) & \mbox{for}& \mu \geq 0 \\\ R_{assimilatory} \times (1-S_{at}) + R_{maintenance} \times (1-B_{at}) & \mbox{for} & \mu < 0\end{array}\right.,$ (S28)

$\frac{d^{14}S}{dt} = -U \times S_{at},$ (S29)

$\frac{d^{14}G}{dt} = A_m \times (f \times S_{at}-\frac{G}{G_m}\times G_{at}),$(S30)

$\frac{d^{14}B}{dt} = \left\{ \begin{array}{rcl}\mu \times B \times G_{at} & \mbox{for}& \mu \geq 0 \\\ \mu \times B \times (1+g) \times B_{at} & \mbox{for} & \mu < 0\end{array}\right.,$ (S31)

$\frac{d^{14}CO_2}{dt} = \left\{ \begin{array}{rcl} R_{assimilatory} \times S_{at} + (R_{maintenance}+R_{growth}) \times G_{at} & \mbox{for}& \mu \geq 0 \\\ R_{assimilatory} \times S_{at} + R_{maintenance} \times B_{at} & \mbox{for} & \mu < 0\end{array}\right..$ (S32)

#### 1.2.2. Monod model

Isotope-specific Monod model is defined using same principles as defined for sub-microbial model. The model is thus defined by following differential equations:

$\frac{d^{12}S}{dt}=-uptake \times (1-S_{at})+death\times (1-MB_{at}),$ (S33)

$\frac{d^{12}MB}{dt}=growth \times (1-S_{at})-death \times (1-MB_{at}),$ (S34)

$\frac{d^{12}CO_2}{dt}=R_{growth} \times (1-S_{at}),$ (S35)

$\frac{d^{14}S}{dt}=-uptake \times S_{at}+death\times MB_{at},$ (S36)

$\frac{d^{14}MB}{dt}=growth \times S_{at}-death \times MB_{at},$ (S37)

$\frac{d^{14}CO_2}{dt}=R_{growth} \times S_{at}.$ (S38)

In eqs. S33 - S38, $MB_{at}$ is defined as:

$MB_{at} = \frac{{}^{14}MB}{{}^{14}MB + {}^{12}MB}.$ (S39)

#### 1.2.3. Pirt model

Isotope-specific Pirt model is defined by following differential equations:

$\frac{d^{12}S}{dt}=-uptake \times (1-S_{at})+death\times (1-MB_{at}),$ (S40)

$\frac{d^{12}MB}{dt}=growth \times (1-S_{at})-death \times (1-MB_{at}) -R_{maintenance} \times (1-MB_{at}),$ (S41)

$\frac{d^{12}CO_2}{dt}=R_{growth} \times (1-S_{at}) + R_{maintenance} \times (1-MB_{at}),$ (S42)

$\frac{d^{14}S}{dt}=-uptake \times S_{at}+death\times MB_{at},$ (S43)

$\frac{d^{14}MB}{dt}=growth \times S_{at}-death \times MB_{at} -R_{maintenance} \times MB_{at},$ (S44)

$\frac{d^{14}CO_2}{dt}=R_{growth} \times S_{at} + R_{maintenance} \times MB_{at}.$ (S45)

Notice, that no isotope discrimination is assumed in all models.

## 2. Individual studies

### 2.1. Glanville et al. (2016)

```{r Glanville, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S1**: Correspondence between the model fit (solid lines) and observations reported in the study of Glanville et al. (2016). **A:** radiocative carbon in chloroform-labile organic carbon over time following either glucose (grey circles) or alanin (empty circle) addition. **B:** radiocative carbon in produced $$CO_2$$ over time following either glucose (grey circles) or alanin (empty circle) addition. **C:** conversion factor between chloroform-labile organic carbon and total soil microbial biomass estimated as a residual fraction of radioactive carbon in microbial biomass that is not released by chloroform vapors. See the main text section 2.1.1. Mathematical model for details about the model."}
#Observations
GlData = read.csv("../SoilMBVariabilityData/Glanville2016.csv")
#Simulations
GlSim <- read.csv("../Manuscript/figure_data/SimGlanville.csv") 

grid.arrange(
  ggplot(GlData, aes(Time, Flush)) + geom_point(cex=6, pch=21, aes(fill = Substrate)) +
  theme_min + ylab(expression(paste("CLC (pmo", l^{14},C~g(DW)^{-1}, ")"))) + xlab("Time (days)") + 
  theme(legend.position = c(0.6, 0.8), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) +
  geom_line(data = GlSim, aes(Time, Flush, color = Substrate), lwd = 1.2) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("A)"),
  ggplot(GlData, aes(Time, CO2)) + geom_point(cex=6, pch=21, aes(fill = Substrate), show.legend = F) +
  theme_min + ylab(expression(paste(CO[2], " (pmo", l^{14},C~g(DW)^{-1}, ")"))) + xlab("Time (days)") + 
  theme(legend.position = c(0.8, 0.4), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) +
  geom_line(data = GlSim, aes(Time, CO2, color = Substrate), lwd = 1.2, show.legend = F) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("B)"),
  ggplot(GlData, aes(Time, kec)) + geom_point(cex=6, pch=21, aes(fill = Substrate), show.legend = F) +
  theme_min + ylab(expression(paste(italic(k[ec])~factor))) + xlab("Time (days)") + 
  theme(legend.position = c(0.8, 0.6), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) + scale_y_continuous(limits = c(0.2, 0.43)) +
  geom_line(data = GlSim, aes(Time, kec, color = Substrate), lwd = 1.2, show.legend = F) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("C)"),
  nrow = 2
)
```

### 2.2. Sparling et al. (1988, 1990)

```{r Sparling, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S2**: Correspondence between the model fit and observations across 13 different soils reported in the studies of Sparling et al. (1988, 1990). Four measured variables are divided into four boxes. Solid line represents 1:1 correspondence between observations and model predictions. Notice that different boxes have different y and x axis scales."}
#Observations
SData = read.csv("../SoilMBVariabilityData/Sparling1990.csv")
#Simulations
SSim <- read.csv("../Manuscript/figure_data/SimSparling.csv") 

my_labeller <- as_labeller(c(CO2="{}^{14}~CO[2]", Flush="{}^{14}~CLC", kec="k[ec]", S="{}^{14}~S"),
                           default = label_parsed)

ggplot(SSim, aes(Measured, Predicted)) + geom_point(cex = 6, pch = 21, fill = "grey") +
  theme_min + facet_wrap(~Variable, scales = "free", labeller = my_labeller) +
  geom_abline(intercept = 0, slope = 1) +
  xlab("Observations") + ylab("Sub-microbial model predictions")

```

### 2.3. Šantrůčková et al. (2004)

```{r Hasan, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S3**: Correspondence between the model fit (solid lines) and observations reported in the studies of Šantrůčková et al. (2004). Four measured variables are divided into four boxes (A - D). Notice that different boxes have different y and x axis scales."}
#Observations
HPData = read.csv("../SoilMBVariabilityData/Santruckova2004.csv")
#Simulations
HPSim <- read.csv("../Manuscript/figure_data/SimHasan.csv") 

grid.arrange(
##Substrate concentration
ggplot(HPData, aes(Time, S)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{14},"S (", mu, "mol(C) " , g^{-1}, "(DW))"))) + 
  xlab("Time (days)") +
  geom_line(data = HPSim, aes(Time, S), lwd = 1.2) +
  ggtitle("A)"),
##Cmic 14
ggplot(HPData, aes(Time, Flush)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{14}," CLC (", mu, "mol(C) " , g^{-1}, "(DW))"))) + 
  xlab("Time (days)") +
  geom_line(data = HPSim, aes(Time, Flush), lwd = 1.2) +
  ggtitle("B)"),
##Cumulative respiration
ggplot(HPData, aes(Time, CO2)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{14},"C", O[2]," (", mu, "mol(C) " , g^{-1}, "(DW))"))) + 
  xlab("Time (days)") +
  geom_line(data = HPSim, aes(Time, CO2), lwd = 1.2) +
  ggtitle("C)"),
##kec factor
ggplot(HPData, aes(Time, kec)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(italic(k[ec])))) + xlab("Time (days)") + #ylim(0.20, 0.35) +
  geom_line(data = HPSim, aes(Time, kec), lwd = 1.2) +
  scale_y_continuous(limits = c(0.2, 0.5)) +
  ggtitle("D)"),
nrow = 2
)

```

### 2.4. Bremer and van Kessel (1990)

```{r Bremer, echo=FALSE, fig.height=12, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S4**: Correspondence between the model fit (solid lines) and observations reported in the study of Bremer and van Kessel (1990). Four measured variables are divided into subplots boxes (A - D). Each subplot is composed of three different boxes, which represent three different treatment - HC - high glucose addition, HCHN - high glucose and high $(NH_{4})_{2} SO_{4}$, LC - low glucose addition. The model parameters are estimated across the treatments. The predictions are not shown for LC treatment due to low correspondence between measured and predicted $k_{ec}$. Notice that different boxes have different y and x axis scales."}
#Observations
BKData = read.csv("../SoilMBVariabilityData/BremerKessel1990.csv")
#Simulations
BKSim <- read.csv("../Manuscript/figure_data/SimBremer.csv") 

grid.arrange(
  ##Substrate concentration
ggplot(BKData, aes(Time, S)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{14},"S (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = BKSim, aes(Time, S), lwd = 1.5) +
  facet_wrap(~Treatment, scales="free") + ggtitle("A)"),
ggplot(BKData, aes(Time, Cmic14)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{14}, "CLC (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = BKSim, aes(Time, Cmic14), lwd = 1.5) +
  facet_wrap(~Treatment, scales="free") + ggtitle("B)"),
##Cumulative respiration
ggplot(BKData, aes(Time, CO214cumul)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + 
  ylab(expression(paste({}^{14},"C", O[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = BKSim, aes(Time, CO2), lwd = 1.5) +
  facet_wrap(~Treatment, scales="free") + ggtitle("C)"),
#kec factor
ggplot(BKData, aes(Time, kec)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(italic(k[ec])))) + xlab("Time (days)") +
  geom_line(data = BKSim, aes(Time, kec), lwd = 1.5) +
  facet_grid(.~Treatment) + ggtitle("D)"),
nrow = 4
)
```

### 2.5. Ziegler et al. (2005)

```{r Ziegler, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S5**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Ziegler et al. (2005). Three measured variables are divided into three subplots boxes (A - C). Notice that different plots have different y and x axis scales."}

ZData = read.csv("../SoilMBVariabilityData/Ziegler2005.csv")
Zsim = read.csv("../Manuscript/figure_data/SimZiegler.csv")
ZsimP = read.csv("../Manuscript/figure_data/SimZieglerP.csv")
ZsimM = read.csv("../Manuscript/figure_data/SimZieglerM.csv")

grid.arrange(
  ##Substrate concentration
ggplot(ZData, aes(Time, S)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{13},"S (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Zsim, aes(Time, S), lwd = 1.5) +
  geom_line(data = ZsimP, aes(Time, S), lwd = 1.5, color = "grey") +
  geom_line(data = ZsimM, aes(Time, S), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("A)"),
##PLFA
ggplot(ZData, aes(Time, PLFA)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{13}~C[PLFA], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Zsim, aes(Time, PLFA), lwd = 1.5) +
  geom_line(data = ZsimP, aes(Time, PLFA), lwd = 1.5, color = "grey") +
  geom_line(data = ZsimM, aes(Time, PLFA), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("B)"),
##Cumulative respiration
ggplot(ZData, aes(Time, CO2cumul)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste({}^{13}~CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Zsim, aes(Time, CO2), lwd = 1.5) +
  geom_line(data = ZsimP, aes(Time, CO2), lwd = 1.5, color = "grey") +
  geom_line(data = ZsimM, aes(Time, CO2), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("C)"),
nrow = 2
)

```

### 2.6. Marstorp and Witter (1999)

```{r Marstorp, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S6**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Martsorp and Witter (1999). Four measured variables are divided into four subplots (A - D). Notice that different plots have different y and x axis scales."}

MData = read.csv("../SoilMBVariabilityData/Marstorp1999.csv")
Msim = read.csv("../Manuscript/figure_data/SimMarstorp.csv")
MsimP = read.csv("../Manuscript/figure_data/SimMarstorpP.csv")
MsimM = read.csv("../Manuscript/figure_data/SimMarstorpM.csv")

grid.arrange(
  ##Substrate concentration
ggplot(MData, aes(Time, S)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("S (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Msim, aes(Time, S), lwd = 1.5) +
  geom_line(data = MsimP, aes(Time, S), lwd = 1.5, color = "grey") +
  geom_line(data = MsimM, aes(Time, S), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("A)"),
##Cmic
ggplot(MData, aes(Time, Flush)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("CLC (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  ylim(0, 15) + geom_line(data = Msim, aes(Time, Flush), lwd = 1.5) +
  geom_line(data = MsimP, aes(Time, Flush), lwd = 1.5, color = "grey") +
  geom_line(data = MsimM, aes(Time, Flush), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("B)"),
##DNA
ggplot(MData, aes(Time, DNA)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("DNA (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Msim, aes(Time, DNA), lwd = 1.5) +
  geom_line(data = MsimP, aes(Time, DNA), lwd = 1.5, color = "grey") +
  geom_line(data = MsimM, aes(Time, DNA), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("C)"),
##Cumulative respiration
ggplot(MData, aes(Time, CO212cumul)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Msim, aes(Time, CO2), lwd = 1.5) +
  geom_line(data = MsimP, aes(Time, CO2), lwd = 1.5, color = "grey") +
  geom_line(data = MsimM, aes(Time, CO2), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("D)"),
nrow = 2
)

```

### 2.7. Tsai et al. (1997)

```{r Tsai, echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S7**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Tsai et al. (1997). Three measured variables are divided into three subplots (A - C). Each subplot is divided into two boxes representing two different treatments - HG - high glucose addition, and LG - low glucose addition. The model parameters are estimated across the treatments. Notice that different plots have different y and x axis scales."}

TData = read.csv("../SoilMBVariabilityData/Tsai1997.csv")
Tsim = read.csv("../Manuscript/figure_data/SimTsai.csv")
TsimP = read.csv("../Manuscript/figure_data/SimTsaiP.csv")
TsimM = read.csv("../Manuscript/figure_data/SimTsaiM.csv")

grid.arrange(
  ##Substrate concentration
ggplot(TData, aes(Time, Cmic)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("CLC (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Tsim, aes(Time, Flush), lwd = 1.5) +
  geom_line(data = TsimP, aes(Time, Flush), lwd = 1.5, color = "grey") +
  geom_line(data = TsimM, aes(Time, Flush), lwd = 1.5, color = "grey30", lty = 2) +
  facet_wrap(~Treatment) +
  ggtitle("A)"),
##Cmic
ggplot(TData, aes(Time, ATP*1000)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("ATP (nmol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Tsim, aes(Time, ATP*1000), lwd = 1.5) +
  geom_line(data = TsimP, aes(Time, ATP*1000), lwd = 1.5, color = "grey") +
  geom_line(data = TsimM, aes(Time, ATP*1000), lwd = 1.5, color = "grey30", lty = 2) +facet_wrap(~Treatment) +
  ggtitle("B)"),
##Cumulative respiration
ggplot(TData, aes(Time, CO2cumul)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Tsim, aes(Time, CO2), lwd = 1.5) +
  geom_line(data = TsimP, aes(Time, CO2), lwd = 1.5, color = "grey") +
  geom_line(data = TsimM, aes(Time, CO2), lwd = 1.5, color = "grey30", lty = 2) +
  facet_wrap(~Treatment) +
  ggtitle("C)"),
nrow = 3
)


```

### 2.8. Joergensen and Raubuch (2002)

```{r Joergensen, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S8**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Joergensen and Raubuch (2002). Three measured variables are divided into three subplots (A - C). Notice that different plots have different y and x axis scales."}

JData = read.csv("../SoilMBVariabilityData/Joergensen2002.csv")
Jsim = read.csv("../Manuscript/figure_data/SimJoerg.csv")
JsimP = read.csv("../Manuscript/figure_data/SimJoergP.csv")
JsimM = read.csv("../Manuscript/figure_data/SimJoergM.csv")

grid.arrange(
  ##Substrate concentration
ggplot(JData, aes(Time, S)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Jsim, aes(Time, S), lwd = 1.5) +
  geom_line(data = JsimP, aes(Time, S), lwd = 1.5, color = "grey") +
  geom_line(data = JsimM, aes(Time, S), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("A)"),
  ##CLC
ggplot(JData, aes(Time, Cmic)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("CLC (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Jsim, aes(Time, Flush), lwd = 1.5) +
  geom_line(data = JsimP, aes(Time, Flush), lwd = 1.5, color = "grey") +
  geom_line(data = JsimM, aes(Time, Flush), lwd = 1.5, color = "grey30", lty = 2) +
  ggtitle("B)"),
##Cmic
ggplot(JData, aes(Time, ATP*1000)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("ATP (nmol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Jsim, aes(Time, ATP*1000), lwd = 1.5) +
  geom_line(data = JsimP, aes(Time, ATP*1000), lwd = 1.5, color = "grey") +
  geom_line(data = JsimM, aes(Time, ATP*1000), lwd = 1.5, color = "grey30", lty = 2)  +  ggtitle("C)"),
nrow = 2
)

```

### 2.9. Nannipieri et al. (1977)

```{r Nannipieri, echo=FALSE, fig.height=10, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S9**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Nannipieri et al. (1977). Three measured variables are divided into three subplots (A - C). Each subplot is further divided into three boxes representing three different treatments - A - high glucose addition with N and P; B - high glucose addition with N; C - low glucose addition. The model parameters are estimated across the treatments. Notice that different plots have different y and x axis scales."}

NData = read.csv("../SoilMBVariabilityData/Nannipieri1977.csv")
Nsim = read.csv("../Manuscript/figure_data/SimNannipieri.csv")
NsimP = read.csv("../Manuscript/figure_data/SimNannipieriP.csv")
NsimM = read.csv("../Manuscript/figure_data/SimNannipieriM.csv")

grid.arrange(
  ##Substrate concentration
ggplot(NData, aes(Time, W/1000)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("Biomass (mg ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Nsim, aes(Time, W/1000), lwd = 1.5) +
  geom_line(data = NsimP, aes(Time, W/1000), lwd = 1.5, color = "grey") +
  geom_line(data = NsimM, aes(Time, W/1000), lwd = 1.5, color = "grey30", lty = 2) +
  facet_wrap(~Treatment) +
  ggtitle("A)"),
  ##ATP
ggplot(NData, aes(Time, ATP*1000)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("ATP (nmol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Nsim, aes(Time, ATP*1000), lwd = 1.5) +
  geom_line(data = NsimP, aes(Time, ATP*1000), lwd = 1.5, color = "grey") +
  geom_line(data = NsimM, aes(Time, ATP*1000), lwd = 1.5, color = "grey30", lty = 2) + facet_wrap(~Treatment) +
  ggtitle("B)"),
##CO2
ggplot(NData, aes(Time, CO2)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Nsim, aes(Time, CO2), lwd = 1.5) +
  geom_line(data = NsimP, aes(Time, CO2), lwd = 1.5, color = "grey") +
  geom_line(data = NsimM, aes(Time, CO2), lwd = 1.5, color = "grey30", lty = 2)  + 
  facet_wrap(~Treatment) + ggtitle("C)"),
nrow = 3
)
```

### 2.10. Blagodatskaya et al. (2014)

```{r Blagodatskaya, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S10**: Correspondence between the fit of three microbially-explicit models (Sub-microbial - solid black line, Monod - dark grey dashed line, and Pirt - light grey solid line) and observations reported in the study of Blagodatskaya et al. (2014). Two measured variables are divided into two subplots (A - B). Each subplot is further divided into two boxes representing two different soils - Rhizosphere and bulk soil. The model parameters are estimated across the soils. Notice that different plots have different y and x axis scales."}

BData = read.csv("../SoilMBVariabilityData/Blagodatskaya2014.csv")
Bsim = read.csv("../Manuscript/figure_data/SimBlag.csv")
BsimP = read.csv("../Manuscript/figure_data/SimBlagP.csv")
BsimM = read.csv("../Manuscript/figure_data/SimBlagM.csv")

grid.arrange(
  ##DNA
ggplot(BData, aes(Time, DNA)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste("DNA (nmol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Bsim, aes(Time, DNA), lwd = 1.5) +
  geom_line(data = BsimP, aes(Time, DNA), lwd = 1.5, color = "grey") +
  geom_line(data = BsimM, aes(Time, DNA), lwd = 1.5, color = "grey30", lty = 2) +
  facet_wrap(~Treatment) +
  ggtitle("A)"),
  ##CO2
ggplot(BData, aes(Time, CO2)) + geom_point(cex=6, pch=21, fill = "grey") +
  theme_min + ylab(expression(paste(CO[2], " (", mu, "mol(C) ", g^{-1}, "(DW))"))) + xlab("Time (days)") +
  geom_line(data = Bsim, aes(Time, CO2), lwd = 1.5) +
  geom_line(data = BsimP, aes(Time, CO2), lwd = 1.5, color = "grey") +
  geom_line(data = BsimM, aes(Time, CO2), lwd = 1.5, color = "grey30", lty = 2)  + 
  facet_wrap(~Treatment) + ggtitle("B)"),
nrow = 2
)

```

## 3. Supplementary tables

**Table S1:** Parameters of mathematical model described in section 2.1.1. Mathematical model in the main text that were used to fit data shown on Fig. 1A.

| Parameter                                 | Value                 | Unit                         |
|--------------------------------|------------------|----------------------|
| Basal biomass (B)                         | 0.14                  | mg                           |
| z ($\frac{I_m \times y_A}{G_m}$)          | 1.25                  | unitless                     |
| Growth costs (g)                          | 0.21                  | $mg(C)~mg(B)^{-1}$           |
| Specific maintenance rate coefficient (m) | $7.53 \times 10^{-8}$ | $mg(C)~mg(B)^{-1}~hour^{-1}$ |
| Relative amount of DNA in pool B ($i_B$)  | 0.063                 | unitless                     |
| Relative amount of RNA in pool G          | 0.21                  | unitless                     |
| Relative amount of RNA in pool B          | 0.058                 | unitless                     |
| Relative amount of proteins in pool G     | 0.53                  | unitless                     |
| Relative amount of proteins in pool B     | 0.64                  | unitless                     |

**Table S2:** Lower and upper bounds of the parameters (see Tab. 1 for explanation of symbols) that have been used for calibration of microbial-explicit model recognizing two pools of biomass (see sections 2.1.1. Mathematical model and 2.3. Model calibration and statistical analyses for details).

| Parameter | Lower limit        | Upper limit       |
|-----------|--------------------|-------------------|
| $I_m$     | 0.01               | 20                |
| $K_m$     | 0.1                | 3000              |
| $y_A$     | 0                  | 1                 |
| $G_m$     | $1 \times 10^{-3}$ | $1 \times 10^{3}$ |
| m         | $1 \times 10^{-8}$ | 1                 |
| g         | 0.01               | 10                |
| $n_G$     | 0                  | 1                 |
| $n_B$     | 0                  | 1                 |
| $i_B$     | 0                  | 1                 |
| $t_G$     | 0                  | 1                 |
| $t_B$     | 0                  | 1                 |
| $p_G$     | 0                  | 1                 |
| $p_B$     | 0                  | 1                 |

## 4. Supplementary figures

```{r Figure1, out.height = "600px", out.width='900px', echo=F, message=FALSE, fig.cap="**Figure S11:** Prisma flow diagram of the publications selection procedure."}
knitr::include_graphics("../Manuscript/prismaMBData.pdf")
```
