---
title: "Understanding the variability in macromolecular composition of soil microbial biomass"
subtitle: "Supplementary Information"
author: "Petr Čapek, Michal Choma, Eva Kaštovská, Karolina Tahovská, Helen Glanville, Hana Šantrůčková"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and ggplot theme, echo=FALSE, warning=FALSE, message=FALSE}
#Libraries
library(ggplot2)
library(dplyr)
library(reshape)
library(minpack.lm)
library(bbmle)
library(deSolve)
library(FME)
library(DEoptim)
library(ABCoptim)
library(rcompanion)
library(optimx)
library(reticulate)
library(gridExtra)
#ggplot theme
theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
                 axis.title=element_text(size=18, colour="black"),
                 axis.line=element_line(size=0.5, colour="black"),
                 strip.text=element_text(size=18, face="bold"),
                 axis.ticks=element_line(size=1, colour="black"),
                 #axis.ticks.length=unit(-0.05, "cm"),
                 panel.background=element_rect(colour="black", fill="white"),
                 panel.grid=element_line(linetype=0),
                 legend.text=element_text(size=14, colour="black"),
                 legend.title=element_text(size=14, colour="black"),
                 legend.position=c("right"),
                 legend.key.size=unit(1, "cm"),
                 strip.background=element_rect(fill="grey98", colour="black"),
                 legend.key=element_rect(fill="white", size=1.2),
                 legend.spacing=unit(0.5, "cm"),
                 plot.title=element_text(size=18, face="bold", hjust=-0.05))
```

## 1. Microbially-explicit models

### 1.1. Natural abundance of carbon isotopes

#### 1.1.1. Monod model

Monod model defines three pools of organic carbon (C) - organic substrate (S), microbial biomass (MB) and carbon dioxide ($CO_2$). The changes of these pools over time are defined by following equations:

$\frac{dS}{dt}=-uptake+death,$ (S1)

$\frac{dMB}{dt}=growth-death,$ (S2)

$\frac{dCO_2}{dt}=R_{growth}.$ (S3)

Microbial biomass consumes organic substrate ($uptake$) by a rate proportional to $MB$ and follows the hyperbolic scaling with $S$:

$uptake = I_m \times MB \times f;~~~f = \frac{S}{K_m + S}.$ (S4)

In eq. S4, $I_m$ is maximum uptake rate constant and $K_m$ is half-saturation constant. Organic substrate taken up is directly converted to $MB$ that has invariable chemical composition. The conversion of $S$ to $MB$ is denoted as $growth$. Growth requires energy. The energy requirements are covered directly from substrate and cause growth associated $CO_2$ production (i.e. $R_{growth}$). Growth and $R_{growth}$ are directly linked to $uptake$ via the parameter $y_A$, which defines the efficiency of organic carbon conversion to $MB$ (i.e. commonly denoted as carbon use efficiency):

$growth = uptake \times y_A,$ (S5)

$R_{growth} = uptake \times (1-y_A).$ (S6)

Along with the growth, $MB$ is constantly dying. The dying part of microbial community is consumed by the growing part of microbial community. The death rate follows the first order decay kinetic:

$death = MB \times d.$ (S7)

In eq. S7, $d$ is the death rate constant.

Since the $MB$ is not directly measured, invariable conversion factors between $MB$ and different proxy-parameters are defined as follows:

$CLC = MB \times k_{ec},$ (S8)

$DNA = MB \times k_{DNA},$ (S9)

$ATP = MB \times k_{ATP},$ (S10)

$PLFA = MB \times k_{PLFA}.$ (S11)

In eqs. S8 - S11, symbols $CLC$, $k_{ec}$, $k_{DNA}$, $k_{ATP}$, and $k_{PLFA}$ denote chloroform-labile organic carbon, and conversion factors between $CLC$, DNA, ATP and PLFA, respectively.

#### 1.1.2. Pirt model

Pirt model is almost identical to Monod model. The only difference is the presence of maintenance. $MB$ has to maintain its integrity. Integrity maintenance requires energy, which is covered from $MB$ and results in $CO_2$ production ($R_{maintenance}$). $R_{maintenance}$ is proportional to $MB$ via the specific maintenance rate constant ($m$):

$R_{maintenance} = MB \times m.$ (S12)

The Pirt model is thus defined by following differential equations:

$\frac{dS}{dt}=-uptake+death,$ (S13)

$\frac{dMB}{dt}=growth-death-R_{maintenance},$ (S14)

$\frac{dCO_2}{dt}=R_{growth}+R_{maintenance}.$ (S15)

Conversion factors between different proxy-parameters and $MB$ is defined the same way as in Monod model (eqs. S8 - S11).

The Monod, Pirt and two-pool biomass model defined in main text represent gradient of increasing complexity and number of model parameters.

### 1.2. Isotope-specific models

#### 1.2.1. Sub-microbial model

The isotope-specific sub-microbial model have eight pools represented by symbols ${}^{12}S$, ${}^{12}B$, ${}^{12}G$, ${}^{12}CO_{2}$, ${}^{14}S$, ${}^{14}B$, ${}^{14}G$ and ${}^{14}CO_{2}$. The left superscripts of the symbols denote either natural carbon isotope 12 or radioactive isotope 14. Sum of two isotopes in respective pool defines the total pool:

$S = {}^{12}S + {}^{14}S,$ (S16)

$B = {}^{12}B + {}^{14}B,$ (S17)

$G = {}^{12}G + {}^{14}G,$ (S18)

$CO_2 = {}^{12}CO_2 + {}^{14}CO_2.$ (S19)

The fluxes between the total pools are defined the same way as described in main text (section 2.1. Mathematical model). Isotope-specific fluxes are proportional to total fluxes by the relative abundance of the source pools. The relative abundance is defined in respect to radioactive carbon isotope as:

$S_{at} = \frac{{}^{14}S}{{}^{14}S + {}^{12}S},$ (S20)

$B_{at} = \frac{{}^{14}B}{{}^{14}B + {}^{12}B},$ (S21)

$G_{at} = \frac{{}^{14}G}{{}^{14}G + {}^{12}G},$ (S22)

$CO_{2at} = \frac{{}^{14}CO_2}{{}^{14}CO_2 + {}^{12}CO_2}.$ (S23)

Implementing eqs. S20 - S21 into eqs. 10 - 13 reported in main text, isotope-specific sub-microbial model is defined as:

$\frac{d^{12}S}{dt} = -U \times (1-S_{at}),$ (S24)

$\frac{d^{12}G}{dt} = A_m \times [f \times (1-S_{at})-\frac{G}{G_m}\times (1-G_{at})],$ (S25)

$\frac{d^{12}B}{dt} = \left\{ \begin{array}{rcl}\mu \times B \times (1-G_{at}) & \mbox{for}& \mu \geq 0 \\\ \mu \times B \times (1+g) \times (1-B_{at}) & \mbox{for} & \mu < 0\end{array}\right.,$ (S26)

$\frac{d^{12}CO_2}{dt} = \left\{ \begin{array}{rcl} R_{assimilatory} \times (1-S_{at}) + (R_{maintenance}+R_{growth}) \times (1-G_{at}) & \mbox{for}& \mu \geq 0 \\\ R_{assimilatory} \times (1-S_{at}) + R_{maintenance} \times (1-B_{at}) & \mbox{for} & \mu < 0\end{array}\right.,$ (S27)

$\frac{d^{14}S}{dt} = -U \times S_{at},$ (S28)

$\frac{d^{14}G}{dt} = A_m \times (f \times S_{at}-\frac{G}{G_m}\times G_{at}),$(S29)

$\frac{d^{14}B}{dt} = \left\{ \begin{array}{rcl}\mu \times B \times G_{at} & \mbox{for}& \mu \geq 0 \\\ \mu \times B \times (1+g) \times B_{at} & \mbox{for} & \mu < 0\end{array}\right.,$ (S30)

$\frac{d^{14}CO_2}{dt} = \left\{ \begin{array}{rcl} R_{assimilatory} \times S_{at} + (R_{maintenance}+R_{growth}) \times G_{at} & \mbox{for}& \mu \geq 0 \\\ R_{assimilatory} \times S_{at} + R_{maintenance} \times B_{at} & \mbox{for} & \mu < 0\end{array}\right..$ (S31)

#### 1.2.2. Monod model

Isotope-specific Monod model is defined using same principles as defined for sub-microbial model. The model is thus defined by following differential equations:

$\frac{d^{12}S}{dt}=-uptake \times (1-S_{at})+death\times (1-MB_{at}),$ (S32)

$\frac{d^{12}MB}{dt}=growth \times (1-S_{at})-death \times (1-MB_{at}),$ (S33)

$\frac{d^{12}CO_2}{dt}=R_{growth} \times (1-S_{at}),$ (S34)

$\frac{d^{14}S}{dt}=-uptake \times S_{at}+death\times MB_{at},$ (S35)

$\frac{d^{14}MB}{dt}=growth \times S_{at}-death \times MB_{at},$ (S36)

$\frac{d^{14}CO_2}{dt}=R_{growth} \times S_{at}.$ (S37)

In eqs. S32 - S37, $MB_{at}$ is defined as:

$MB_{at} = \frac{{}^{14}MB}{{}^{14}MB + {}^{12}MB}.$ (S38)

#### 1.2.3. Pirt model

Isotope-specific Pirt model is defined by following differential equations:

$\frac{d^{12}S}{dt}=-uptake \times (1-S_{at})+death\times (1-MB_{at}),$ (S39)

$\frac{d^{12}MB}{dt}=growth \times (1-S_{at})-death \times (1-MB_{at}) -R_{maintenance} \times (1-MB_{at}),$ (S40)

$\frac{d^{12}CO_2}{dt}=R_{growth} \times (1-S_{at}) + R_{maintenance} \times (1-MB_{at}),$ (S41)

$\frac{d^{14}S}{dt}=-uptake \times S_{at}+death\times MB_{at},$ (S42)

$\frac{d^{14}MB}{dt}=growth \times S_{at}-death \times MB_{at} -R_{maintenance} \times MB_{at},$ (S43)

$\frac{d^{14}CO_2}{dt}=R_{growth} \times S_{at} + R_{maintenance} \times MB_{at}.$ (S43)

Notice, that no isotope discrimination is assumed in all models.

## 2. Individual studies

### 2.1. Glanville et al. (2016)

```{r Glanville, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S1**: Correspondence between the model fit (solid lines) and observations reported in the study of Glanville et al. (2016). **A:** radiocative carbon in chloroform-labile organic carbon over time following either glucose (grey circles) or alanin (empty circle) addition. **B:** radiocative carbon in produced $$CO_2$$ over time following either glucose (grey circles) or alanin (empty circle) addition. **C:** conversion factor between chloroform-labile organic carbon and total soil microbial biomass estimated as a residual fraction of radioactive carbon in microbial biomass that is not released by chloroform vapors. See the main text section 2.1.1. Mathematical model for details about the model."}
#Observations
GlData = read.csv("../SoilMBVariabilityData/Glanville2016.csv")
#Simulations
GlSim <- read.csv("../Manuscript/figure_data/SimGlanville.csv") 

grid.arrange(
  ggplot(GlData, aes(Time, Flush)) + geom_point(cex=6, pch=21, aes(fill = Substrate)) +
  theme_min + ylab(expression(paste("CLC (pmo", l^{14},C~g(DW)^{-1}, ")"))) + xlab("Time (days)") + 
  theme(legend.position = c(0.6, 0.8), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) +
  geom_line(data = GlSim, aes(Time, Flush, color = Substrate), lwd = 1.2) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("A)"),
  ggplot(GlData, aes(Time, CO2)) + geom_point(cex=6, pch=21, aes(fill = Substrate), show.legend = F) +
  theme_min + ylab(expression(paste(CO[2], " (pmo", l^{14},C~g(DW)^{-1}, ")"))) + xlab("Time (days)") + 
  theme(legend.position = c(0.8, 0.4), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) +
  geom_line(data = GlSim, aes(Time, CO2, color = Substrate), lwd = 1.2, show.legend = F) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("B)"),
  ggplot(GlData, aes(Time, kec)) + geom_point(cex=6, pch=21, aes(fill = Substrate), show.legend = F) +
  theme_min + ylab(expression(paste(italic(k[ec])~factor))) + xlab("Time (days)") + 
  theme(legend.position = c(0.8, 0.6), legend.title = element_blank()) + 
  scale_fill_manual(values = c("white", "grey")) + scale_y_continuous(limits = c(0.2, 0.43)) +
  geom_line(data = GlSim, aes(Time, kec, color = Substrate), lwd = 1.2, show.legend = F) +
  scale_color_manual(values = c("grey30", "black")) + ggtitle("C)"),
  nrow = 2
)
```

### 2.2. Sparling et al. (1988, 1990)

```{r Sparling, echo=FALSE, fig.height=8, fig.width=8, message=FALSE, warning=FALSE, fig.cap="**Figure S1**: Correspondence between the model fit (solid lines) and observations reported in the study of Glanville et al. (2016). **A:** radiocative carbon in chloroform-labile organic carbon over time following either glucose (grey circles) or alanin (empty circle) addition. **B:** radiocative carbon in produced $$CO_2$$ over time following either glucose (grey circles) or alanin (empty circle) addition. **C:** conversion factor between chloroform-labile organic carbon and total soil microbial biomass estimated as a residual fraction of radioactive carbon in microbial biomass that is not released by chloroform vapors. See the main text section 2.1.1. Mathematical model for details about the model."}
#Observations
SData = read.csv("../SoilMBVariabilityData/Sparling1990.csv")
#Simulations
SSim <- read.csv("../Manuscript/figure_data/SimSparling.csv") 



```

## 3. Supplementary tables

**Table S1:** Parameters of mathematical model described in section 2.1.1. Mathematical model in the main text that were used to fit data shown on Fig. 1A.

| Parameter                                 | Value                 | Unit                         |
|--------------------------------|------------------|----------------------|
| Basal biomass (B)                         | 0.14                  | mg                           |
| z ($\frac{I_m \times y_A}{G_m}$)          | 1.25                  | unitless                     |
| Growth costs (g)                          | 0.21                  | $mg(C)~mg(B)^{-1}$           |
| Specific maintenance rate coefficient (m) | $7.53 \times 10^{-8}$ | $mg(C)~mg(B)^{-1}~hour^{-1}$ |
| Relative amount of DNA in pool B ($i_B$)  | 0.063                 | unitless                     |
| Relative amount of RNA in pool G          | 0.21                  | unitless                     |
| Relative amount of RNA in pool B          | 0.058                 | unitless                     |
| Relative amount of proteins in pool G     | 0.53                  | unitless                     |
| Relative amount of proteins in pool B     | 0.64                  | unitless                     |

**Table S2:** Lower and upper bounds of the parameters (see Tab. 1 for explanation of symbols) that have been used for calibration of microbial-explicit model recognizing two pools of biomass (see sections 2.1.1. Mathematical model and 2.3. Model calibration and statistical analyses for details).

| Parameter | Lower limit        | Upper limit       |
|-----------|--------------------|-------------------|
| $I_m$     | 0.01               | 20                |
| $K_m$     | 0.1                | 3000              |
| $y_A$     | 0                  | 1                 |
| $G_m$     | $1 \times 10^{-3}$ | $1 \times 10^{3}$ |
| m         | $1 \times 10^{-8}$ | 1                 |
| g         | 0.01               | 10                |
| $n_G$     | 0                  | 1                 |
| $n_B$     | 0                  | 1                 |
| $i_B$     | 0                  | 1                 |
| $t_G$     | 0                  | 1                 |
| $t_B$     | 0                  | 1                 |
| $p_G$     | 0                  | 1                 |
| $p_B$     | 0                  | 1                 |

## 4. Supplementary figures

```{r Figure1, out.height = "600px", out.width='900px', echo=F, message=FALSE, fig.cap="**Figure S1:** Prisma flow diagram of the publications selection procedure."}
knitr::include_graphics("../Manuscript/prismaMBData.pdf")
```
