---
title: "Understanding the variability in macromolecular composition of soil microbial biomass"
author: "Petr Čapek, Michal Choma, Eva Kaštovská, Karolina Tahovská, Helen Glanville, Hana Šantrůčková"
date: "`r Sys.Date()`"
output: word_document
bibliography: references.bib
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries and ggplot theme, echo=FALSE, warning=FALSE, message=FALSE}
#Libraries
library(ggplot2)
library(dplyr)
library(reshape)
library(minpack.lm)
library(bbmle)
library(deSolve)
library(FME)
library(DEoptim)
library(ABCoptim)
library(rcompanion)
library(optimx)
library(reticulate)
library(gridExtra)
library(ggpomological)
#ggplot theme
theme_min<-theme(axis.text.x=element_text(vjust=0.2, size=18, colour="black"),
                 axis.text.y=element_text(hjust=0.2, size=18, colour="black"),
                 axis.title=element_text(size=18, colour="black"),
                 axis.line=element_line(size=0.5, colour="black"),
                 strip.text=element_text(size=18, face="bold"),
                 axis.ticks=element_line(size=1, colour="black"),
                 #axis.ticks.length=unit(-0.05, "cm"),
                 panel.background=element_rect(colour="black", fill="white"),
                 panel.grid=element_line(linetype=0),
                 legend.text=element_text(size=14, colour="black"),
                 legend.title=element_text(size=14, colour="black"),
                 legend.position=c("right"),
                 legend.key.size=unit(1, "cm"),
                 strip.background=element_rect(fill="grey98", colour="black"),
                 legend.key=element_rect(fill="white", size=1.2),
                 legend.spacing=unit(0.5, "cm"),
                 plot.title=element_text(size=18, face="bold", hjust=-0.05))

```

## 1. Introduction

Macromolecular and elemental composition of cells of single-species microbial communities is known to vary significantly [@Hanegraaf2001; @Makino2003; @herbert1961]. As quoted by Denis Herbert, "Such a statement as 'the ribonucleic acid content of Bacillus cereus is 16.2%' is by itself as incomplete and misleading as the statement 'the boiling-point of water is 70°C'.". Variability in cell composition is primarily associated with changes in specific growth rate (Fig. 1A, @Hanegraaf2001; @Herbert1961; @Makino2003). To grow fast, microbial cells require several macromolecules and metabolites with unique elemental composition [@Vrede2004; @Franklin2011] that are transformed into various cell components or perform the transformation. Typical example of macromolecule mediating growth is ribosomal RNA, which translates genetic code into a sequence of amino acids. Because the rate of translation per unit of one ribosome is constant, increase in growth rate can be achieved only by increasing the ribosomal RNA content [@Vrede2004; @Franklin2011]. For that reason, ribosomal RNA content typically increases with increasing specific growth rate (Fig. 1A, @Hanegraaf2001; @Makino2003; @Herbert1961).

To describe the relationship between the growth rate and cell composition, it proved convenient to distinguish at least two pools of microbial biomass (Fig. 1C, @Hanegraaf2001; @Franklin2011). One pool represents all cell constituents that are present in non-growing cell. The second pool represents the macromolecules and metabolites that are associated with increased specific growth rate [@Hanegraaf2001; @Franklin2011]. The terminology is not unified in respect to these pools because they are abstract. [@Hanegraaf2001] for example, denote them as "Structures" and "Reserves", respectively. [@Franklin2011] use the terms "baseline biomass" and "growth machinery" (authors also distinguish the third pool of "uptake machinery"). For convenience, we will hereafter use the terms "Basal biomass" ($B$) and "Growth intermediates" ($G$). As the $G$ increases, specific growth rate increases accordingly (see section 2.1. for detail description). Defining the macromolecular content and/or elemental stoichiometry of pools $B$ and $G$ allows then to associate changes in cell composition directly to changes in specific growth rate (Fig. 1A; @Hanegraaf2001; @Hanegraaf2000). The pools $B$ and $G$ represent a reliable simplification of otherwise complex microbial physiology.

Soil microbial communities are characteristic by co-occurrence of many species in the same microenvironments. Given the fact that different species might have different characteristics such as basal DNA or RNA content, it remains unclear if any generalizable relationship between the specific growth rate and composition of soil microbial biomass can exist. Several empirical evidences suggest that it can. By tracking the radiolabelled glucose added to soil at different concentrations, @bremer1990 found that two pools of soil microbial biomass should be distinguished to explain variability in chloroform-labile content of microbial biomass, i.e. the $k_{ec}$ factor. The similar conclusion was later on reached by @nguyen2001 and @glanville2016. Other studies showed that the ratio between chloroform-labile organic carbon (CLC) and various macromolecules (i.e. DNA @Marstorp1999; @malik2015, ATP @tsai1997; @joergensen2002, RNA and PLFA @malik2015; @gunina2017) changes significantly and consistently over time or across organic substrate (typically glucose) concentrations. @tessier1998 found a consistent relationship between the total soil organic carbon (C) content, which affects the growth rate of microbial biomass, and the $k_{ec}$ factor. Recently, @hagerty2022 showed that considering two pools of soil microbial biomass significantly improves a simulation accuracy of microbially-explicit mathematical models.

If the generalizable relationship between the microbial biomass composition and the specific-growth rate can be defined for soil microbial communities, it would have four important implications that are intimately linked. The first two implications are methodological. *(i)* Understanding the relationship between the microbial biomass composition and the specific growth rate can improve the quantification of the total soil microbial biomass. Most of the methods relies on quantitative extraction of some part of microbial biomass, so called biomass proxy parameter (e.g. CLC @Vance1987, DNA @Marstorp1999, PLFA @ziegler2005, ATP @blagodatskaya2013). Total microbial biomass is then calculated as a product of the proxy parameter and arbitrary defined conversion factor, which is, in many cases, assumed to be invariant even though the empirical results prove the opposite. For example, the $k_{ec}$ factor changes significantly over the course of few days after organic substrate addition [@glanville2016; @bremer1990; @santruckova2004] and may vary by a factor of up to 30 among different soils or soil depths [@sparling1989; @sparling1988; @Dictor1998]. The specific growth rate could explain at least part of this variability [@tessier1998; @bremer1990]. *(ii)* Improved quantification of total soil microbial biomass improves the calculation of biomass-specific physiological parameters, such as carbon use efficiency [@geyer2019; @hagerty2022] and turnover rate [@chen2019], which are based on the flux analysis of the isotopically labelled substrates through the total microbial biomass. *(iii)* When two biomass proxies show consistently distinct relationship to specific growth rate, their ratio can be used as a simple indicator of the growth rate. (iv) At last, improved mechanistic and quantitative understanding of soil microbial community physiology may improve the simulation accuracy of microbially-explicit soil biogeochemical models even above the level calculated by @hagerty2022, whose analysis sticks to the assumption of constant composition of soil microbial biomass.

The aim of this study is to define the relationship between the macromolecular composition of soil microbial biomass, and to determine its validity and generalizability. To do so, we first attempt to explain the temporal variability in $k_{ec}$ factor and ratios between different macromolecules individually as reported in previously published studies. We use a microbially-explicit mathematical model inspired by the Dynamic Energy Budget theory [@Hanegraaf2001; @kooijman1991; @Kooijman2009]. Then, we use this model to explain the variability in all data products across all published studies using one set of model parameters. We hypothesize that defining the two pools of soil microbial biomass and their macromolecular content can explain the variability in $k_{ec}$ factor as well as ratios of different macromolecules within and across different soils.

## 2. Material and Methods

### 2.1. Theory

#### 2.1.1. Mathematical model

Here we present a theoretical background and formulation of a mathematical model, which is inspired by the Dynamic Energy Budget theory [@Hanegraaf2001; @kooijman1991; @Kooijman2009], and is used to explain variability in macromolecular composition of soil microbial biomass. The model is hereafter denoted as "sub-microbial". As compared to Dynamic Energy Budget theory, the sub-microbial model is defined directly on C-1 molar basis to avoid excessive number of conversion factors, and the terminology common to soil microbial studies is used. For simplicity, we assume that the microbial community is composed of heterotrophic microorganisms, volume and shape of all microbial cells do not change over time and it is similar across soils, the aerobic conditions dominate in all microenvironments, the catabolic reactions within the cell degrades organic substrate completely to $CO_{2}$ (i.e. release of fermentation products to soil is negligible), and the extracellular enzymes production is negligible (see the data selection for detail explanation).

The sub-microbial model is schematically shown on Fig. 1C (see also Tab. 1 for list of symbols and their explanation) and recognizes four different carbon pools - organic substrate ($S$), $B$, $G$, and $CO_{2}$. Total microbial biomass ($MB$) is a sum of $B$ and $B \times G$ because $G$ is scaled to $B$ (i.e. growth intermediates are located inside the cell). $S$ is the entire source of energy and C for microbial biomass. Its uptake rate ($U$) is proportional to $B$ (i.e. $G$ has no effect) and follows classical hyperbolic scaling with $S$:

$U = I_m \times B \times f;~~~f = \frac{S}{K_m + S}.$ (1)

In eq. 1, $I_m$ is maximum uptake rate of a substrate and $K_M$ is half-saturation constant. Following uptake, $S$ gets assimilated, i.e. converted to $G$. There are overhead energy costs of assimilation resulting in production of $CO_{2}$ (Fig. 1C, assimilatory respiration). It is assumed that these costs are covered directly from $S$. Substrate assimilation ($A$) and assimilatory respiration ($R_{assimilatory}$) rates are respectively defined as:

$A = U \times y_A,$ (2)

and

$R_{assimilatory} = U \times (1-y_A).$ (3)

In eqs. 2 and 3, $y_A$ defines the efficiency of substrate conversion into $G$. Once the $S$ is assimilated, $G$ can be mobilized and used to either cover maintenance costs or/and to grow (i.e. produce $B$). The mobilization rate of $G$($J$) is defined as:

$J = A_m \times \frac{G}{G_m}.$ (4)

In eq. 4, $A_m$ is specific maximum assimilation rate constant, which is equal to $I_m \times y_A$. $G_m$ is maximum content of $G$ that the cell can have. Thus, the more substrate the cell assimilates into growth intermediates, the faster the $G$ can be mobilized. $G$ content cannot exceed $G_m$, so $\frac{G}{G_m}$ is always a number between 0 and 1 (see Fig. 1B). When $\frac{G}{G_m}$ is 1, $J$ reaches the maximum under the given environmental conditions.

At any time, microbial cell has to maintain integrity of its basal cellular material $B$. $G$ requires no maintenance. Cellular integrity maintenance is associated with $CO_{2}$ production because it requires energy (Fig. 1C, maintenance respiration). The maintenance respiration ($R_{maintenance}$) is defined as:

$R_{maintenance} = m \times B.$ (5)

In eq. 5, $m$ is the specific maintenance rate coefficient.

If the amount of $G$ is sufficiently high (see the details below), microbial biomass can grow. Growth is simply defined as a change of $B$ over time ($\frac{dB}{dt}$). Because the growth is defined as a conversion of $G$ to $B$, it is associated with energy costs (Fig. 1C, growth respiration) and thus, growth respiration ($R_{growth}$) as defined by eq. 6:

$R_{growth} = g \times \frac{dB}{dt},$ (6)

in which $g$ defines the energy cost of growth in terms of amount of $CO_{2}$ produced per unit of produced $B$.

Maintenance of cellular integrity has an absolute priority over growth. Thus, growth is realized only if $J$ is higher than $m$. Fluxes $R_{maintenance}$, $R_{growth}$ and $\frac{dB}{dt}$ balance the change of $G$ over time:

$\underbrace{m \times B}_{R_{maintenance}} + \underbrace{g \times \frac{dB}{dt}}_{R_{growth}} + \underbrace{\frac{dB}{dt}}_{growth} = \underbrace{J \times B}_{G~mobilization~rate~within~cells} - \underbrace{\frac{dB}{dt} \times G}_{dilution~of~G~by~growth}.$ (7)

The second term on the right hand side of eq. 7 defines the dilution of pool $G$ by the growth. Because $G$ is scaled to $B$, an increase of pool $B$ causes a decrease of $G$. The last term on the right hand side ensures that $B \times G$ remains constant. Without that term, the mass balance would be violated. Eq. 7 can be rearranged to equation 8 to define a formulation of $\frac{dB}{dt}$ at growing conditions:

$\frac{dB}{dt} = \mu \times B;~~\mu = \frac{J-m}{1+g+G}.$ (8)

Eq. 8 shows that the specific growth rate ($\mu$) is a function of $J$, $m$, $g$ and $G$ - a variables, which are all scaled to pool $B$ (i.e. biomass specific). It also shows that $\mu$ is positive when the amount of $G$ is high enough to produce flux $J$ exceeding $m$. In opposite case, $\mu$ becomes negative. If $G$ is so small that it cannot cover the maintenance costs, these costs must be covered from $B$. The eq. 8, however, must be rearranged to:

$\frac{dB}{dt} = \mu \times B \times (1+g).$ (9)

because $\mu \times (1+g)$ simplifies to $-m$ when $G$ is zero and the eq. 9 becomes complementary to eq. 5. This formulation aligns well with the empirical observations. In nature, the microbial cells may temporarily maintain its basal metabolism to survive the starvation period when growth ceases. During this time, the cell shrinks (and/or changes shape) because all maintenance C costs are met from the pool of basal cellular material (i.e. depolymerized cellular structures, @mason1993). The period of basal metabolism maintenance can last up to several months in some cases [@morita1993]. For that reason, the death rate is not defined here directly. It is assumed that loss of microbial biomass is only due to maintenance respiration in short term (see data selection for details). Another reason for omitting the death rate formulation is to keep number of model parameters as low as possible so the model remains parsimonious and reasonable parameter estimates can be made [@Sierra2015].

To summarize the theoretical background, the mathematical model depicted on Fig. 1C is described by four differential equations defined respectively for pools $S$, $G$, $B$, and $CO_2$ as follows:

$\frac{dS}{dt} = -U,$ (10)

$\frac{dG}{dt} = A_m \times (f-\frac{G}{G_m}),$ (11)

$\frac{dB}{dt} = \left\{ \begin{array}{rcl}\mu \times B & \mbox{for}& \mu \geq 0 \\\ \mu \times B \times (1+g) & \mbox{for} & \mu < 0\end{array}\right.,$ (12)

$\frac{dCO_2}{dt} = R_{assimilatory} + R_{maintenance}+min(R_{growth}, 0).$ (13)

The last term of eq. 13 ($min(R_{growth}, 0)$) means that the growth respiration occurs only when growth is realized ($\mu$ is positive).

When the model is specified for the experiments supplying carbon substrates labelled by heavy or radioactive isotope to soil, each pool needs to be divided into two that correspond to two different C isotopes. Flows of two different isotopes follow the same rules and their ratio is defined by the relative amount of radioactive isotope of the source pool. The detail formulation of isotope-specific sub-microbial models is reported in Supplementary Information, section 1.2. Isotope-specific models.

#### 2.1.2. Conversion factors

To define the relationship between specific growth rate (eq. 8) and macromolecular composition of soil microbial biomass, conversion factors need to be specified. A macromolecule, e.g. RNA, could belong to either pool $G$, pool $B$ or to both pools in different proportions (@Hanegraaf2001, Tab. S1). When soil microbial community is exposed to chloroform vapor, some amount of $G$ and $B$ is released to soil and extracted. $k_{ec}$ factor can be thus defined as the weighted mean of extractable parts of $G$ and $B$:

$k_{ec} = \frac{n_G \times G \times B + n_B \times B}{G \times B + B} = \frac{n_G \times G + n_B}{G + 1}.$ (14)

In. eq. 14, $n_G$ and $n_B$ represent the relative amounts of pools $G$ and $B$, respectively, which can be extracted from soil after chloroform fumigation. The relationship between the specific growth rate $\mu$ and soil microbial biomass composition (in respect to $k_{ec}$ in this case) derives from combination of equations 8 and 14. $\mu$ as well as $k_{ec}$ directly depend on $G$. The higher the amount of growth intermediates the cell has, the faster they can be mobilized (i.e. the higher the flux $J$, eq. 4) and used to achieve high specific growth rate (unless $m$ is extremely high for any reason, eq. 8). The maximum specific growth rate is bounded by the maximum amount of growth intermediates the cell can has ($G_m$). If $n_G$ and $n_B$ are distinctly different, $k_{ec}$ changes along the gradient of specific growth rate (Fig. 1B). Fig. 1B shows $k_{ec}$ resulting from different combinations of both parameters along the gradient of $\frac{G}{G_m}$. The higher the difference between the $n_G$ and $n_B$, the higher the $k_{ec}$ variability along the gradient. If the $n_G$ is higher than $n_B$, $k_{ec}$ increases with increasing specific growth rate and vice versa. In the published studies, $k_{ec}$ tends to decrease over time following the glucose addition [@bremer1990; @santruckova2004; @glanville2016]. If this trend results from the growth rate deceleration after glucose depletion, it would suggest that $n_G$ is higher than $n_B$.

Eq. 14 can be rewritten to define a conversion factor between any macromolecule or biomass proxy and total soil microbial biomass. Here we define the conversion factors respectively for DNA ($k_{DNA}$), ATP ($k_{ATP}$), and PLFA ($k_{PLFA}$) as follows:

$k_{DNA} = \frac{i_B}{G + 1},$ (15)

$k_{ATP} = \frac{t_G \times G + t_B}{G + 1},$ (16)

$k_{PLFA} = \frac{p_B}{G + 1}.$ (17)

According to @Hanegraaf2001, DNA is only part of pool $B$ so Eq. 15 contains only parameter $i_{B}$ defining the relative abundance of DNA in $B$. Because PLFA are produced only to form cell membranes, we assume they belong only to pool $B$ similarly as DNA.

In addition, @nannipieri1978 reports mass of microbial biomass ($W$) along with its ATP content. The total mass of microbial community results from the sum of masses of pools $B$ and $G$ denoted as $W_B$ and $W_G$, respectively.

### 2.2. Data collection

To test the accuracy of the theoretical background, the model defined above was confronted with data published in the literature. Because the aim of the study is to analyze the relationship between the growth rate and composition of soil microbial biomass, we searched for studies in which the growth rate or composition of microbial biomass was reported to change. The studies were searched on the Web of Science, PubMed and Scopus (Fig. S11) using the keywords 'soil', 'microbial biomass', 'growth', 'carbon', 'ATP', 'DNA', 'PLFA', 'radiolabel', 'isotope', 'substrate addition', and their combinations. This search resulted in 3924 indexed studies. We also searched in citations of four most relevant publications [@bremer1990; @glanville2016; @Marstorp1999; @nguyen2001] resulting in additional (partially overlapping) 343 studies. These studies were analyzed and filtered according to following criteria: (i) the selected experiments had to be short-term to make sure the death rate and extracellular enzyme production can be neglected in mass balance, (ii) the data had to be available for more than three different time points, (iii) at least three different C pools or variables had to be measured and ideally two of them are different biomass proxies, and (iv) initial conditions had to be well defined (Fig. S11, i.e. note that several studies did not comply with more than one criteria). Applying these criteria, only seven eligible studies were selected and further analyzed @bremer1990; @glanville2016; @joergensen2002; @Marstorp1999; @santruckova2004; @tsai1997; @nannipieri1978. Due to the low number of eligible studies, four exceptions were made. Study of @blagodatskaya2014 was selected regardless the criterion (iii) to increase the number of studies reporting change of DNA concentration over time following saturating concentration of organic substrate. Study of @ziegler2005 was again selected regardless the criterion (iii) because it is the only study reporting change of PLFA concentration over time with sufficient temporal resolution. Studies of @sparling1988 and @Sparling1990a were selected regardless the criterion (ii) because they report large number of $k_{ec}$ estimates for soils with reasonably well defined initial conditions.

Data were extracted from publications either directly or using the WebPlotDigitizer. When some data points were overlapping in graphs, they were not extracted. When suitable, all extracted data were recalculated to $\mu mol$ of carbon per gram of oven dried soil (i.e. except of e.g. mass of microbial community) applying conversion factors reported in @Vrede2004. When total microbial biomass was reported, the respective biomass proxy, which was directly measured, was calculated back using a conversion factor reported in the study (i.e. typically 0.45 for $k_{ec}$ factor). The data are available at <https://github.com/petacapek/SoilMBVariability>.

### 2.3. Model calibration and statistical analyses

The model was first calibrated against the data reported in individual studies to asses its ability to explain the temporal variability in different C pools and $k_{ec}$ factor following substrate addition. Model parameters were adjusted to minimize the objective function $J$ using the Artificial Bee Colony algorithm [@vegayon2017] within a predefined range of parameters' values (Tab. S2). The objective function $J$ was defined for each measured variable as:

$J = \sum\limits_{i=1}^{n}{(\frac{P_{i}-O_{i}}{SD})^{2}} ,$ (18)

where $O_{i}$ and $P_{i}$ stand for observation $i$ and its corresponding value predicted by the model, and $SD$ is the standard deviation of all observations. The objective functions calculated for each measured variable were summed up. Uncertainty of parameters estimates were additionally determined by Constrained Markov Chain Monte Carlo simulation on 5000 iterations using R package FME [@Soetaert2010] applying the parameter values defined by the Artificial Bee Colony algorithm as initial starting point. To evaluate the goodness of correspondence between model predictions and data, coefficient of determination $(R^{2})$ was calculated.

We further evaluated if the goodness of Sub-microbial model fit is significantly better than the fit of two most commonly used microbial-explicit models of Monod and Pirt, which assume constant composition of soil microbial biomass (and therefore have less parameters). Both models are described in Supplementary Information, section 1. Microbially-explicit models. The significance of the difference was evaluated across all measured data (i.e. all measured variables at all time points). Because different variables have different means and variability, all data were first standardized to have zero mean and unit variance. This was achieved by subtracting means of respective variables from individual observations and dividing by standard deviation. Sum of squared differences between the standardized observations and standardized predictions provides normalized residual sum of squares ($SS_{res}^{normalized}$ ), which is independent of variable identity (i.e. $CO_2$ production or ATP concentration). Using this value, F test was performed to evaluated the significance of goodness of fit difference between the models. The F value was calculated as:

$F = \frac{(Model_{SS_{res}^{normalized}}^{1} - Model_{SS_{res}^{normalized}}^{2}) \times (N_{obs}-N_{par}^{2})}{Model_{SS_{res}^{normalized}}^{2} \times (N_{par}^{2}-N_{par}^{1})}.$ (19)

In eq. 19, $Model_{SS_{res}^{normalized}}^{1}$ denotes ${SS_{res}^{normalized}}$ calculated for Monod or Pirt model. $Model_{SS_{res}^{normalized}}^{2}$ then denotes ${SS_{res}^{normalized}}$ calculated for Sub-microbial model. $N_{obs}$ is number of all observations (including all measured variables), $N_{par}^{1}$ is number of model parameters of either Monod or Pirt models, and $N_{par}^{2}$ is number of parameters of Sub-microbial model.

The comparison was not performed in studies reporting directly measured $k_{ec}$ factor, against which the Sub-microbial model was calibrated. Monod and Pirt models cannot predict the $k_{ec}$ variability so they cannot be calibrated against it and the comparison with Sub-microbial model is thus meaningless.

To further asses generalizability of the theoretical background, Sub-microbial model calibration was performed across all studies and the goodness of model fit was analyzed. To minimize number of model parameters, study of @ziegler2005 was excluded from the analysis and only ATP content from study of @nannipieri1978 was used. Thus, the conversion factors between total soil microbial biomass, CFC, DNA, and ATP were estimated in global analysis (see Tab. 1). Due to the low temporal resolution, studies of @sparling1988 and @Sparling1990a (see above) were also excluded. Studies supplying isotopically labelled and unlabelled organic substrate were analyzed together (see the discussion section).

To perform calibration of model parameters, initial conditions must be defined. This is not always possible when total soil biomass is composed of two pools whose content is apriori unknown. In studies that report temporal dynamic of two microbial proxies, the initial pools $B$ and $G$ were calculated as a product of conversion factors between total soil microbial biomass and reported proxies, and the initial concentration of the proxies. In other studies, $G$ was assumed to be initially equal to $\frac{m \times G_{m}}{I_m \times y_A}$ so the $G$ is large enough to only cover maintenance costs. When the model was calibrated against data from all studies combined, we simply assumed that initial $G$ is close to zero.

The mathematical models were written and run in python [@vanrossum1995]. Parameter estimation and all statistical analyses were performed in R [@rcoreteam2020].

## 3. Results

### 3.1. Individual studies

The correspondence between model predictions and observations in each individual study is graphically shown in Supplementary Information section 2. Individual studies. The summary of the results is reported in Table 2. In table 3, model parameters optimized to fit the data within individual studies are reported.

Sub-microbial model was able to explain large amount of variability in published data, often exceeding 70%. The best performance was achieved against data of @santruckova2004 (Fig. 2A). The sub-microbial model explained almost all the variability in temporal trend of glucose concentration, ${}^{14}CLC$, ${}^{14}CO_2$, and $k_{ec}$ following glucose addition using eight parameters (Tab. 3). In four studies that determined the $k_{ec}$, sub-microbial model explained over 70% of variability even when calibrated across different treatments (e.g. @bremer1994, Fig. S4). Nevertheless, the model was not able to reproduce the temporal trend of $k_{ec}$ in treatment supplying the lowest amount of glucose (see Fig. S4). In studies of @sparling1988 and @Sparling1990a, 14% of variability in $k_{ec}$ was explained by the model. It must be noted, however, that this amount of explained variability was derived across 13 different soils.

Sub-microbial model was able to explain the changes in ratios between various microbial biomass proxies over time. The example for DNA to CLC ratio [@Marstorp1999], and ATP to CLC ratio [@joergensen2002] are visualized on Fig. 2C and D, respectively. Monod and Pirt models predicted constant ratio between the proxies (Fig. 2C, D) and thus, both fitted the data significantly worse than Sub-microbial model (Tab. 4). Sub-microbial model did not fit the data reported in study of @ziegler2005 better than Monod and Pirt models. In this study, PLFA was reported as the only biomass proxy.

In respect to conversion factors, both $n_G$ and $n_B$ estimates defining the $k_{ec}$ factor were relatively similar across studies (Tab. 3). Two estimates of parameter $i_B$ defining the DNA concentration in basal pool of microbial biomass were almost the same (Tab. 3). The greatest variability was observed in parameter $t_G$, which defines the ATP content in pool of growth intermediates. Whereas the $t_G$ was estimated to be higher than $t_B$ in study of @tsai1997, its was much lower in studies of @joergensen2002 and @nannipieri1978 (Tab. 3).

### 3.2. Global fit of Sub-microbial model

When calibrated across all studies combined, Sub-microbial model explained significant amount of variability (over 50%) in most of the individual C pools (Tab. 3 and Fig. 3). Most of the variability was explained in DNA and substrate concentrations (93% and 80%, respectively). Lowest amount of variability, on the other hand, was explained in isotopically labelled $CO_2$. The sub-microbial model tended to overestimate ${}^{14}CO_2$ and underestimate ${}^{14}CLC$ at the same time (Fig. 3). 34% of variability in $k_{ec}$ factor was explained across all studies (Tab. 3).

Conversion factors $n_G$ and $n_B$ were estimated to be 0.42 and 0.24, respectively (Tab. 3). These values make the $k_{ec}$ factor vary between 0.24 and 0.36. Conversion factor $i_B$ was 0.052 and the conversion factors $t_B$ and $t_G$ were almost indistinguishable from each other.

## 4. Discussion

## Tables

**Table 1:** List of symbols used in mathematical model formulation. Explanation and units are reported to each symbol. Parameters marked by symbol '\*' and '$\ddagger$' are subject of calibration performed in individual studies and across studies, respectively (see section 2.3. Model calibration and statistical analyses for details).

|      **Symbol**       |                           **Explanation**                            |                    **Unit**                     |
|:----------------:|:-----------------------------:|:---------------------:|
|       **Pools**       |                                                                      |                                                 |
|           S           |                          Organic substrate                           |             $\mu mol(C)~g(DW)^{-1}$             |
|           B           |                            Basal biomass                             |             $\mu mol(C)~g(DW)^{-1}$             |
|           G           |                         Growth intermediates                         |     $\mu mol(C)~\mu mol(C~of~pool~B)^{-1}$      |
|        $CO_2$         |                            Carbon dioxide                            |             $\mu mol(C)~g(DW)^{-1}$             |
|          MB           |            Total microbial biomass ($MB=(B+B \times G)$)             |             $\mu mol(C)~g(DW)^{-1}$             |
|           W           |                  Total mass of microbial community                   |               $\mu g~g(DW)^{-1}$                |
|      **Fluxes**       |                                                                      |                                                 |
|           U           |                Organic substrate uptake rate (eq. 1)                 |        $\mu mol(C)~g(DW)^{-1}~day^{-1}$         |
|           A           |        Rate of organic substrate assimilation into G (eq. 1)         |        $\mu mol(C)~g(DW)^{-1}~day^{-1}$         |
|  $R_{assimilatory}$   |                   Assimilatory respiration (eq. 3)                   |        $\mu mol(C)~g(DW)^{-1}~day^{-1}$         |
|           J           |                    Mobilization rate of G (eq. 4)                    | $\mu mol(C)~\mu mol(C~of~pool~B)^{-1} day^{-1}$ |
|   $R_{maintenance}$   |                   Maintenance respiration (eq. 5)                    |        $\mu mol(C)~g(DW)^{-1}~day^{-1}$         |
|     $R_{growth}$      |                      Growth respiration (eq. 5)                      |        $\mu mol(C)~g(DW)^{-1}~day^{-1}$         |
|         $\mu$         |                     Specific growth rate (eq. 6)                     | $\mu mol(C)~\mu mol(C~of~pool~B)^{-1} day^{-1}$ |
|    **Parameters**     |                                                                      |                                                 |
| ${}^{\ddagger} I_m$\* |                         Maximum uptake rate                          | $\mu mol(C)~\mu mol(C~of~pool~B)^{-1} day^{-1}$ |
| ${}^{\ddagger}K_m$\*  |                       Half-saturation constant                       |             $\mu mol(C)~g(DW)^{-1}$             |
| ${}^{\ddagger}y_A$\*  |                       Assimilation efficiency                        |                    unitless                     |
|         $A_m$         |             Maximum assimilation rate ($I_m \times y_A$)             | $\mu mol(C)~\mu mol(C~of~pool~B)^{-1} day^{-1}$ |
| ${}^{\ddagger}G_m$\*  |                     Maximum amount of G within B                     |     $\mu mol(C)~\mu mol(C~of~pool~B)^{-1}$      |
|  ${}^{\ddagger}$m\*   |                Specific maintenance rate coefficient                 | $\mu mol(C)~\mu mol(C~of~pool~B)^{-1} day^{-1}$ |
|  ${}^{\ddagger}$g\*   |                             Growth costs                             |     $\mu mol(C)~\mu mol(C~of~pool~B)^{-1}$      |
| ${}^{\ddagger}n_G$\*  | Relative amount of G extracted from soil after chloroform fumigation |                    unitless                     |
| ${}^{\ddagger}n_B$\*  | Relative amount of B extracted from soil after chloroform fumigation |                    unitless                     |
| ${}^{\ddagger}i_B$\*  |                   Relative amount of DNA in pool B                   |                    unitless                     |
| ${}^{\ddagger}t_G$\*  |                   Relative amount of ATP in pool G                   |                    unitless                     |
| ${}^{\ddagger}t_B$\*  |                   Relative amount of ATP in pool B                   |                    unitless                     |
|        $p_B$\*        |                  Relative amount of PLFA in pool B                   |                    unitless                     |
|        $W_G$\*        |            C-1 mole to mass conversion factor for pool G             |        $\mu g~\mu mol(C~of~pool~G)^{-1}$        |
|        $W_B$\*        |            C-1 mole to mass conversion factor for pool B             |        $\mu g~\mu mol(C~of~pool~B)^{-1}$        |

**Table 2:** Capability (reported as $R^2$) of sub-microbial model to explain variability in different data products (see Table 1 for explanation of the abbreviations) reported in 11 studies. Studies are divided into two categories - studies supplying either isotope-labelled or unlabelled substrate. Substrate identity is reported in the table. The $R^2$ is reported separately for the model fit to data reported in individual studies and to all data combined together.

| Study                          | Substrate | Variable                 |              |                                |          |                   |      |
|---------|---------|---------|---------|---------|---------|---------|---------|
| Labelled substrate             |           |                          |              |                                |          |                   |      |
|                                |           | ${}^{14}S$ or ${}^{13}S$ | ${}^{14}CLC$ | ${}^{14}CO_2$ or ${}^{13}CO_2$ | $k_{ec}$ | ${}^{13}C_{PLFA}$ |      |
| Glanville et al. (2016)        | Glucose   |                          | 0.85         | 0.99                           | 0.77     |                   |      |
|                                | Alanine   |                          | 0.67         | 0.98                           | 0.69     |                   |      |
| Sparling et al. (1988, 1990)   |           | 0.98                     | 0.84         | 0.78                           | 0.14     |                   |      |
| Šantrůčková et al. (2004)      | Glucose   | 0.98                     | 0.99         | 0.99                           | 0.99     |                   |      |
| Bremer and van Kessel (1990)   | Glucose   | 0.99                     | 0.98         | 0.99                           | 0.76     |                   |      |
| $\dagger$Ziegler et al. (2005) | Glucose   | 0.82                     |              | 0.94                           |          | 0.72              |      |
| $\ddagger$All studies          |           | 0.56                     | 0.57         | 0.22                           | 0.34     |                   |      |
| Unlabelled substrate           |           |                          |              |                                |          |                   |      |
|                                |           | S                        | CLC          | $CO_2$                         | DNA      | ATP               | W    |
| Marstorp and Witter (1999)     | Glucose   | 0.85                     | 0.45         | 0.93                           | 0.86     |                   |      |
| Tsai et al. (1997)             | Glucose   |                          | 0.73         | 0.95                           |          | 0.90              |      |
| Joergensen and Raubuch (2002)  | Glucose   | 0.96                     | 0.91         |                                |          | 0.96              |      |
| Nannipieri et al. (1977)       | Glucose   |                          |              | 0.86                           |          | 0.71              | 0.70 |
| Blagodatskaya et al. (2014)    | Glucose   |                          |              | 0.91                           | 0.86     |                   |      |
| $\ddagger$All studies          |           | 0.80                     | 0.65         | 0.84                           | 0.93     | 0.49              |      |

${}^{\dagger}$ ${}^{13}C$ labelled glucose was supplied in this study (radiolabelled substrate was supplied in the others)

${}^{\ddagger}$Studies of Sparling et al. (1988 and 1990) and Ziegler et al. (2005) were excluded from the analysis. See the Methods section for explanation.

**Table 3:** Parameters (see Table 1 for explanation of symbols) of sub-microbial model calibrated either separately against data reported reported in 11 different studies or against all data combined${}^{\ddagger}$. Best parameters and thier standard deviations are reported.

| Study                         | Substrate | $I_m$           | $K_m$           | $y_A$                                         | $G_m$                                         | m                                             | g                              | $n_G$            | $n_{B}$         | $i_B$              | $t_G$                                         | $t_B$                                         | $p_B$                                         | $W_G$           | $W_B$           |
|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|-----|
| Glanville et al. (2016)       | Glucose   | $13.5 \pm 3.37$ | $0.12 \pm 0.22$ | $0.94 \pm 0.05$                               | $6.39 \pm 1.23$                               | $8.15 \times 10^{-4} \pm 3.07 \times 10^{-4}$ | $0.15 \pm 0.09$                | $0.43 \pm 0.03$  | $0.28 \pm 0.02$ |                    |                                               |                                               |                                               |                 |                 |
|                               | Alanine   | $9.86 \pm 3.33$ | $0.11 \pm 0.10$ | $0.89 \pm 0.06$                               | $4.57 \pm 1.61$                               | $1.24 \times 10^{-3} \pm 1.55 \times 10^{-3}$ | $0.29 \pm 0.12$                | $0.36 \pm 0.02$  | $0.27 \pm 0.02$ |                    |                                               |                                               |                                               |                 |                 |
| Sparling et al. (1988, 1990)  | Glucose   | $0.11 \pm 0.04$ | $0.15 \pm 0.13$ | $0.67 \pm 0.05$                               | $0.76 \pm 0.79$                               | $6.21 \times 10^{-7} \pm 5.08 \times 10^{-7}$ | $0.02 \pm 4.58 \times 10^{-3}$ | $0.42 \pm 0.10$  | $0.15 \pm 0.06$ |                    |                                               |                                               |                                               |                 |                 |
| Šantrůčková et al. (2004)     | Glucose   | $11.3 \pm 4.40$ | $212 \pm 64.6$  | $0.98 \pm 0.08$                               | $2.52 \pm 1.12$                               | $1.97 \times 10^{-6} \pm 1.86 \times 10^{-6}$ | $0.33 \pm 0.12$                | $0.48 \pm 0.18$  | $0.22 \pm 0.04$ |                    |                                               |                                               |                                               |                 |                 |
| Bremer and van Kessel (1990)  | Glucose   | $4.00 \pm 2.01$ | $25.0 \pm 18.8$ | $0.98 \pm 0.08$                               | $0.93 \pm 0.95$                               | $4.88 \times 10^{-3} \pm 5.27 \times 10^{-3}$ | $0.51 \pm 0.18$                | $0.48 \pm 0.04$  | $0.34 \pm 0.05$ |                    |                                               |                                               |                                               |                 |                 |
| Ziegler et al. (2005)         | Glucose   | $8.67 \pm 3.74$ | $323 \pm 82.6$  | $0.99 \pm 0.03$                               | $1.60 \times 10^{-3} \pm 1.92 \times 10^{-3}$ | $0.01 \pm 0.004$                              | $0.01 \pm 0.01$                |                  |                 |                    |                                               |                                               | $7.98 \times 10^{-3} \pm 2.20 \times 10^{-3}$ |                 |                 |
| Marstorp and Witter (1999)    | Glucose   | $6.28 \pm 2.57$ | $222 \pm 59.6$  | $1 \pm 0.09$                                  | $1.08 \pm 1.1$                                | $7.45 \times 10^{-3} \pm 4.62 \times 10^{-3}$ | $1.15 \pm 0.45$                | $1.00 \pm 0.16$  | $0.15 \pm 0.07$ | $0.028 \pm 0.009$  |                                               |                                               |                                               |                 |                 |
| Tsai et al. (1997)            | Glucose   | $0.87 \pm 3.54$ | $323 \pm 63.5$  | $0.43 \pm 0.16$                               | $0.84 \pm 1.08$                               | $3.51 \times 10^{-4} \pm 6.60 \times 10^{-3}$ | $0.19 \pm 0.38$                | $0.40 \pm 0.17$  | $0.03 \pm 0.04$ |                    | $1.18 \times 10^{-3} \pm 9.93 \times 10^{-5}$ | $2.71 \times 10^{-5} \pm 9.51 \times 10^{-5}$ |                                               |                 |                 |
| Joergensen and Raubuch (2002) | Glucose   | $20.0 \pm 4.70$ | $252 \pm 90.3$  | $0.68 \pm 0.13$                               | $3.37 \pm 1.91$                               | $1.84 \times 10^{-2} \pm 5.57 \times 10^{-3}$ | $5.00 \pm 0.87$                | $0.69 \pm 0.17$  | $0.46 \pm 0.10$ |                    | $1.52 \times 10^{-8}\pm 1.59 \times 10^{-4}$  | $8.61 \times 10^{-4} \pm 1.41 \times 10^{-4}$ |                                               |                 |                 |
| Nannipieri et al. (1977)      | Glucose   | $0.66 \pm 0.67$ | $132 \pm 57.8$  | $1.12 \times 10^{-2} \pm 7.18 \times 10^{-3}$ | $1.12 \times 10^{-3} \pm 3.21 \times 10^{-3}$ | $8.23 \times 10^{-3} \pm 4.77 \times 10^{-3}$ | $0.73 \pm 0.95$                |                  |                 |                    | $1.58 \times 10^{-5} \pm 1.09 \times 10^{-5}$ | $2.28 \times 10^{-4} \pm 2.68 \times 10^{-4}$ |                                               | $8.09 \pm 4.77$ | $19.6 \pm 21.8$ |
| Blagodatskaya et al. (2014)   |           | $2.58 \pm 2.73$ | $0.10 \pm 0.13$ | $1.00 \pm 0.12$                               | $1.21 \pm 1.41$                               | $7.65 \times 10^{-5} \pm 12.8 \times 10^{-5}$ | $0.97 \pm 1.41$                |                  |                 | $0.027 \pm 0.0065$ |                                               |                                               |                                               |                 |                 |
| $\ddagger$All studies         |           | $5.74 \pm 0.26$ | $206 \pm 22.3$  | $0.95 \pm 0.04$                               | $1.99 \pm 0.16$                               | $2.13 \times 10^{-3} \pm 3.50 \times 10^{-4}$ | $1.82 \pm 0.20$                | $0.456 \pm 0.02$ | $0.24 \pm 0.01$ | $0.056 \pm 0.004$  | $2.52 \times 10^{-4} \pm 4.28 \times 10^{-5}$ | $2.96 \times 10^{-4} \pm 4.20 \times 10^{-5}$ |                                               |                 |                 |

${}^{\ddagger}$Studies of Sparling et al. (1988 and 1990) and Ziegler et al. (2005) were excluded from the analysis. See the Methods section for explanation.

**Table 4:** Comparison of goodness of fit of three different microbially-explicit models (Sub-microbial, Monod and Pirt). Normalized (across different data products) residual sum of squares ($SS_{res}^{normalized}$), number of model parameters ($N_{par}$), and number of observations ($N_{obs}$), against which the models are calibrated, are reported in the table. F value expresses the difference between $SS_{res}^{normalized}$ calculated for Sub-microbial model and Monod or Pirt model given the number of model parameters and number of observations (see eq. 19). The positive value means the decrease of $SS_{res}^{normalized}$ (i.e. better model fit). p value expresses the probability that the difference in $SS_{res}^{normalized}$ is accidental.

| Study                         | Model         | $SS_{res}^{normalized}$ | $N_{par}$ | $N_{obs}$ | F     | p       |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
| Ziegler et al. (2005)         | Sub-microbial | 6.51                    | 7         | 16        |       |         |
|                               | Pirt          | 2.82                    | 6         | 16        | -5.10 | 1       |
|                               | Monod         | 3.77                    | 6         | 16        | -1.89 | 1       |
| Marstorp and Witter (1999)    | Sub-microbial | 18.2                    | 9         | 66        |       |         |
|                               | Pirt          | 35.0                    | 7         | 66        | 36.3  | \<0.001 |
|                               | Monod         | 35.3                    | 6         | 66        | 17.8  | \<0.001 |
| Tsai et al. (1997)            | Sub-microbial | 9.78                    | 10        | 72        |       |         |
|                               | Pirt          | 31.8                    | 7         | 72        | 46.5  | \<0.001 |
|                               | Monod         | 33.2                    | 6         | 72        | 37.2  | \<0.001 |
| Joergensen and Raubuch (2002) | Sub-microbial | 0.93                    | 10        | 21        |       |         |
|                               | Pirt          | 7.08                    | 7         | 21        | 23.9  | \<0.001 |
|                               | Monod         | 7.00                    | 6         | 21        | 18.2  | \<0.001 |
| Nannipieri et al. (1977)      | Sub-microbial | 16.2                    | 10        | 68        |       |         |
|                               | Pirt          | 39.1                    | 7         | 68        | 27.2  | \<0.001 |
|                               | Monod         | 20.7                    | 6         | 68        | 4.01  | 0.006   |
| Blagodatskaya et al. (2014)   | Sub-microbial | 5.16                    | 7         | 90        |       |         |
|                               | Pirt          | 7.77                    | 6         | 90        | 42.0  | \<0.001 |
|                               | Monod         | 7.76                    | 5         | 90        | 20.9  | \<0.001 |

## Figures

```{r Figure1, out.height = "700px", out.width='800px', echo=F, message=FALSE, fig.cap="**Figure 1:** **A)** Relationship between the specific growth rate of Aerobacter aerogenes population, its cell mass (dark grey circles) and three different cell components - proteins (empty cirles), RNA (light grey circles), and DNA (black circles). Data were redrawn from Herbert (1961). Lines represent fit of the model described in section 2.1.1. Mathematical model. Model parameters that were used to draw the predictions are reported in Tab. S1. **B)** Relationship between the amount of growth intermediates scaled to the maximum that microbial cell can have $$(\\frac{G}{G_m})$$ and the $$k_{ec}$$ factor (i.e. chloroform labile part of microbial biomass) depending on the value of parameters $$n_G$$ and $$n_B$$ representing respectively the relative amount of basal microbial biomass and growth intermediates extrated from soil after chloroform fumigation (see Tab. 1 and section 2.1.1. Mathematical model for detail explanation). **C)** Schematic representation of sub-microbial model described in section 2.1.1. Mathematical model. Different boxes represent different carbon pools. Arrows represent different carbon fluxes. "}
knitr::include_graphics("../Manuscript/Theoretical.pdf")
```

```{r Figure2, echo=FALSE, fig.height=10, fig.width=10, message=FALSE, warning=FALSE, fig.cap="**Figure 2**: A - B: Correspondence between the predictions of sub-microbial model (black solid lines) and observations reported in Šantrůčková et al. (2004) - $k_{ec}$ factor (grey circles in plot A), ${}^{14}CLC$ (empty circles in plot B), ${}^{14}CO_2$ (black circles in plot B), and ${}^{14}S$ (grey circles in plot B). C: Development of DNA to CLC ratio over time following glucose addition as observed in study of Marstorp and Witter (1999). Solid black line represent predictions of sub-microbial model, solid ligh grey and dashed dark grey lines represent predictions of Pirt and Monod model, respectively. C: Development of ATP to CLC ratio over time following glucose addition as observed in study of Joergensen and Raubuch (2002). Solid black line represent predictions of sub-microbial model, solid ligh grey and dashed dark grey lines represent predictions of Pirt and Monod model, respectively. "}

JData = read.csv("../SoilMBVariabilityData/Joergensen2002.csv")
Jsim = read.csv("../Manuscript/figure_data/SimJoerg.csv")
JsimP = read.csv("../Manuscript/figure_data/SimJoergP.csv")
JsimM = read.csv("../Manuscript/figure_data/SimJoergM.csv")
MData = read.csv("../SoilMBVariabilityData/Marstorp1999.csv")
Msim = read.csv("../Manuscript/figure_data/SimMarstorp.csv")
MsimP = read.csv("../Manuscript/figure_data/SimMarstorpP.csv")
MsimM = read.csv("../Manuscript/figure_data/SimMarstorpM.csv")
HPData = read.csv("../SoilMBVariabilityData/Santruckova2004.csv")
HPSim <- read.csv("../Manuscript/figure_data/SimHasan.csv") 

grid.arrange(
  ggplot(HPData, aes(Time, kec)) + geom_point(cex=6, pch=21, fill = "grey") +
    theme_min + ylab(expression(paste(italic(k[ec])))) + xlab("Time (days)") + 
    geom_line(data = HPSim, aes(Time, kec), lwd = 1.2) +
    scale_y_continuous(limits = c(0.2, 0.5)) +
    ggtitle("A)"),
  ggplot(HPData, aes(x = Time)) + geom_point(cex=6, pch=21, aes(y = S, fill = "Substrate")) +
    geom_point(cex=6, pch=21, aes(y = Flush*4, fill = "CLC")) +
    geom_point(cex=6, pch=21, aes(y = CO2*4, fill = "CO2")) +
    geom_line(data = HPSim, aes(Time, S), lwd = 0.8, show.legend = F) +
    geom_line(data = HPSim, aes(Time, Flush*4), lwd = 0.8, show.legend = F) +
    geom_line(data = HPSim, aes(Time, CO2*4), lwd = 0.8, show.legend = F) +
    scale_y_continuous(sec.axis = sec_axis(~./4,
                                           name = expression(paste({}^{14}," CLC and ",{}^{14}~CO[2] ," (", mu, "mol(C) " , g^{-1}, "(DW))")))) + 
    theme_min + theme(legend.position = c(0.8, 0.3), legend.title = element_blank()) +
    labs(y = expression(paste({}^{14},"S (", mu, "mol(C) " , g^{-1}, "(DW))")),
         x = expression(paste("Time (days)")),
         fill = "Variable", color = "Variable") +
    scale_fill_manual(values = c("white", "black", "grey"), 
                      labels = c(expression(paste({}^{14}~CLC)),
                                 expression(paste({}^{14}~CO[2])),
                                 expression(paste({}^{14}~S))))+
    scale_color_manual(values = c("black", "grey80", "grey"), 
                       labels = c(expression(paste({}^{14}~CLC)),
                                  expression(paste({}^{14}~CO[2])),
                                  expression(paste({}^{14}~S)))) +
    ggtitle("B)"),
 ggplot(MData, aes(Time, DNA/Flush)) + geom_point(cex=6, pch=21, fill = "grey") +
    theme_min + ylab(expression(paste(frac(DNA, CLC)~~~frac(mol,mol)))) + xlab("Time (days)") +
    geom_line(data = Msim, aes(Time, DNA/Flush, color = "Sub-microbial"), lwd = 1.5, show.legend = F) +
    geom_line(data = MsimP, aes(Time, DNA/Flush, color = "Pirt"), lwd = 1.5, show.legend = F) +
    geom_line(data = MsimM, aes(Time, DNA/Flush, color = "Monod"), lwd = 1.5, lty = 2, show.legend = F) +
    ggtitle("C)") +
    scale_color_manual(values = c("grey30", "grey", "black")) +
    labs(color = "Model"),
  ggplot(JData, aes(Time, ATP*1000/Cmic)) + geom_point(cex=6, pch=21, fill = "grey") +
    theme_min + ylab(expression(paste(frac(ATP, CLC)~~~frac(mmol,mol)))) + xlab("Time (days)") +
    geom_line(data = Jsim, aes(Time, ATP*1000/Flush, color = "Sub-microbial"), lwd = 1.5) +
    geom_line(data = JsimP, aes(Time, ATP*1000/Flush, color = "Pirt"), lwd = 1.5) +
    geom_line(data = JsimM, aes(Time, ATP*1000/Flush, color = "Monod"), lwd = 1.5, lty = 2)  +  
    ggtitle("D)") + theme(legend.position = c(0.7, 0.2), legend.title = element_blank()) + ylim(0.6, 2.2) +
    scale_color_manual(values = c("grey30", "grey", "black")) +
    labs(color = "Model")
)

```

```{r Figure3, echo=FALSE, fig.height=12, fig.width=14, message=FALSE, warning=FALSE, fig.cap="**Figure 3**: Correspondence between the predictions of sub-microbial model and observations reported in eight published studies (denoted by different colors). Figure is divided into nine boxes that correspond to nine different data products against which the sub-microbial model was calibrated. Solid black line represents 1:1 correspondence between observations and model predictions. Notice that each box have different y and x-axis scales. "}

SimAll <- read.csv("../Manuscript/figure_data/SimAll.csv")

#Labels and annotation
VizLab <- c('CO2' = "CO[2]", 'DNA' = "DNA", '14CO2' = "atop(14, )~CO[2]", 'kec' = "italic(k[ec])",
            '14MBC' = "atop(14, )~CLC", '14S' = "atop(14, )~S", 'S' = "S", 'ATP' = "ATP",
            'Flush' = "CLC")
VizAnot <- data.frame(Variable = as.character(unique(SimAll$Variable)), label = c("R^{2}==0.84", "R^{2}==0.93", "R^{2}==0.22",
                                                         "R^{2}==0.34", "R^{2}==0.57", "R^{2}==0.56",
                                                         "R^{2}==0.80", "R^{2}==0.49", "R^{2}==0.65"))

ggplot(SimAll, aes(Observations, Predictions)) + geom_point(cex = 6, aes(color = Study)) + theme_min + 
  geom_text(data = VizAnot, mapping = aes(x = -Inf, y = -Inf, label = label),
            hjust = -2.5, vjust = -0.5, parse = T, cex = 6) +
  facet_wrap(~Variable, scales = "free", labeller = as_labeller(VizLab, label_parsed)) + scale_color_pomological() +
  geom_abline(intercept = 0, slope = 1) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

## References:
