---
title: "Understanding the variability in macromolecular composition of soil microbial biomass"
subtitle: "Supplementary Information"
author: "Petr Čapek, Michal Choma, Eva Kaštovská, Karolina Tahovská, Helen Glanville, Hana Šantrůčková"
date: "`r Sys.Date()`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Microbially-explicit models

### 1.1. Monod model

Monod model defines three pools of organic carbon (C) - organic substrate (S), microbial biomass (MB) and carbon dioxide ($CO_2$). The changes of these pools over time are defined by following equations:

$\frac{dS}{dt}=-uptake+death,$ (S1)

$\frac{dMB}{dt}=growth-death,$ (S2)

$\frac{dCO_2}{dt}=R_{growth}.$ (S3)

Microbial biomass consumes organic substrate ($uptake$) by a rate proportional to $MB$ and follows the hyperbolic scaling with $S$:

$uptake = I_m \times MB \times f;~~~f = \frac{S}{K_m + S}.$ (S4)

In eq. S4, $I_m$ is maximum uptake rate constant and $K_m$ is half-saturation constant. Organic substrate taken up is directly converted to $MB$ that has invariable chemical composition. The conversion of $S$ to $MB$ is denoted as $growth$. Growth requires energy. The energy requirements are covered directly from substrate and cause growth associated $CO_2$ production (i.e. $R_{growth}$). Growth and $R_{growth}$ are directly linked to $uptake$ via the parameter $y_A$, which defines the efficiency of organic carbon conversion to $MB$ (i.e. commonly denoted as carbon use efficiency):

$growth = uptake \times y_A,$ (S5)

$R_{growth} = uptake \times (1-y_A).$ (S6)

Along with the growth, $MB$ is constantly dying. The dying part of microbial community is consumed by the growing part of microbial community. The death rate follows the first order decay kinetic:

$death = MB \times d.$ (S7)

In eq. S7, $d$ is the death rate constant.

Since the $MB$ is not directly measured, invariable conversion factors between $MB$ and different proxy-parameters are defined as follows:

$CLC = MB \times k_{ec},$ (S8)

$DNA = MB \times k_{DNA},$ (S9)

$ATP = MB \times k_{ATP},$ (S10)

$PLFA = MB \times k_{PLFA}.$ (S11)

In eqs. S8 - S11, symbols $CLC$, $k_{ec}$, $k_{DNA}$, $k_{ATP}$, and $k_{PLFA}$ denote chloroform-labile organic carbon, and conversion factors between $CLC$, DNA, ATP and PLFA, respectively.

### 1.2. Pirt model

Pirt model is almost identical to Monod model. The only difference is the presence of maintenance. $MB$ has to maintain its integrity. Integrity maintenance requires energy, which is covered from $MB$ and results in $CO_2$ production ($R_{maintenance}$). $R_{maintenance}$ is proportional to $MB$ via the specific maintenance rate constant ($m$):

$R_{maintenance} = MB \times m.$ (S12)

The Pirt model is thus defined by following differential equations:

$\frac{dS}{dt}=-uptake+death,$ (S13)

$\frac{dMB}{dt}=growth-death-R_{maintenance},$ (S14)

$\frac{dCO_2}{dt}=R_{growth}+R_{maintenance}.$ (S15)

Conversion factors between different proxy-parameters and $MB$ is defined the same way as in Monod model (eqs. S8 - S11).

The Monod, Pirt and two-pool biomass model defined in main text represent gradient of increasing complexity and number of model parameters.

## 2. Supplementary tables

**Table S1:** Parameters of mathematical model described in section 2.1.1. Mathematical model in the main text that were used to fit data shown on Fig. 1A.

| Parameter                                 | Value                 | Unit                         |
|--------------------------------|------------------|----------------------|
| Basal biomass (B)                         | 0.14                  | mg                           |
| z ($\frac{I_m \times y_A}{G_m}$)          | 1.25                  | unitless                     |
| Growth costs (g)                          | 0.21                  | $mg(C)~mg(B)^{-1}$           |
| Specific maintenance rate coefficient (m) | $7.53 \times 10^{-8}$ | $mg(C)~mg(B)^{-1}~hour^{-1}$ |
| Relative amount of DNA in pool B ($i_B$)  | 0.063                 | unitless                     |
| Relative amount of RNA in pool G          | 0.21                  | unitless                     |
| Relative amount of RNA in pool B          | 0.058                 | unitless                     |
| Relative amount of proteins in pool G     | 0.53                  | unitless                     |
| Relative amount of proteins in pool B     | 0.64                  | unitless                     |

**Table S2:** Lower and upper bounds of the parameters (see Tab. 1 for explanation of symbols) that have been used for calibration of microbial-explicit model recognizing two pools of biomass (see sections 2.1.1. Mathematical model and 2.3. Model calibration and statistical analyses for details).

| Parameter | Lower limit        | Upper limit       |
|-----------|--------------------|-------------------|
| $I_m$     | 0.01               | 20                |
| $K_m$     | 0.1                | 3000              |
| $y_A$     | 0                  | 1                 |
| $G_m$     | $1 \times 10^{-3}$ | $1 \times 10^{3}$ |
| m         | $1 \times 10^{-8}$ | 1                 |
| g         | 0.01               | 10                |
| $n_G$     | 0                  | 1                 |
| $n_B$     | 0                  | 1                 |
| $i_B$     | 0                  | 1                 |
| $t_G$     | 0                  | 1                 |
| $t_B$     | 0                  | 1                 |
| $p_G$     | 0                  | 1                 |
| $p_B$     | 0                  | 1                 |

## 3. Supplementary figures

```{r Figure1, out.height = "600px", out.width='900px', echo=F, message=FALSE, fig.cap="**Figure S1:** Prisma flow diagram of the publications selection procedure."}
knitr::include_graphics("../Manuscript/prismaMBData.pdf")
```
